---
title: "Individual Assignment (WIP)"
description: |
  Student hard at work. Check back next time. 
draft: TRUE
preview: images/wip.png
author:
  - name: Kelly Koh
    url: https://www.linkedin.com/in/kellykkw/
    affiliation: School of Computing and Information Systems, Singapore Management University
    affiliation_url: https://scis.smu.edu.sg/
date: 06-10-2021
output:
  distill::distill_article:
    toc: true
    toc_depth: 3
    toc_float: true
    self_contained: false
---

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
knitr::opts_chunk$set(fig.retina = 3,
                      include = TRUE,
                      echo = FALSE,
                      eval = TRUE,
                      message = FALSE,
                      warning = FALSE)
```

## 1. Background

### 1.1 Vast Challenge 2021 - The Kronos Incident
In the roughly twenty years that Tethys-based GAStech has been operating a natural gas production site in the island country of Kronos, it has produced remarkable profits and developed strong relationships with the government of Kronos. However, GAStech has not been as successful in demonstrating environmental stewardship.

In January, 2014, the leaders of GAStech are celebrating their new-found fortune as a result of the initial public offering of their very successful company. In the midst of this celebration, several employees of GAStech go missing. An organization known as the Protectors of Kronos (POK) is suspected in the disappearance, but things may not be what they seem.

### 1.2 Mini Challenge 2 Background
GAStech provides many of their employees with company cars for their personal and professional use, but unbeknownst to the employees, the cars are equipped with GPS tracking devices. We are given tracking data for the two weeks leading up to the disappearance, as well as credit card transactions and loyalty card usage data. From this data, we would identify anomalies and suspicious behaviors and identify which people use which credit and loyalty cards.

## 2. Literature Review

### 2.1 Overview on Previous Works
VAST Challenge 2021 made use of the same dataset that was used for VAST Challenge 2014. Of the 28 submissions for the Mini-Challenge 2 in 2014, many of the winning teams made use of D3.js, a JavaScript library to develop their propriety tools. Some of the teams explored novel approach such as using parallel coordinates plot, time rings and network graph. However, most common approaches can be summarized as:

- Timeline of employee’s location color coded by location category overlaid with spend
- Geospatial map of movement
- Space-time cube of movement

Timeline paired with geospatial map was the most effective
The space-time cube was especially hard to interpret. 


### 2.1 Limitations on Previous Works
However, the results cannot be replicated as they are not using open-source software. 


## 3. Methodology and Visualisation Design 

### 3.1 Data Matching To Find Anomalies

One of the challenges of managing the dataset is the lack of definite joining criteria. Fuzzy matching technique has to be employed to match the dataset that do not have the exact data granularity. For example, 
- Loyalty data only has date level information while credit card is has date and time
- Credit card spend are point-in-time while events recognized from GPS data have time windows
- Coordinates of a single location derived from GPS latitude and longitude can vary

The package `fuzzyjoin` of which functions `fuzzy_join` and `geo_join` will be used to match date/spend and lat/long variables.

A methodical approach is taken for data matching. Matching start from the more definite criteria (e.g. exact location, price and date between Credit Card and Loyalty transactions) and removing spurious matches before fuzzy joins are used. The anomalies captured from each fuzzy join will also be discussed. 

### 3.2 Identifying Locations

Once

### 3.3 Visualizing Spend Details Using Interactive Gannt Chart 

### 3.4 Visualizing Event Pattern Using Interative Gannt Chart 

### 3.5 Visualizing GPS Movement Using Interactive Map


## 4. R Packages & Data

These R packages are used:
```{r download packages}
# The following R packages are used for respective parts of the workflow:
packages = c( # Loading data from csv files
               'readr',
              # Cleaning and manipulating data 
              'tidyverse', 'DT', 'fuzzyjoin',
              # Manipulate time and date variables
              'lubridate', 'clock', 
              # Plot heat maps, gantt, parallel coordinates and scatter plot graphs
              'ggplot2', 'igraph', 'gghighlight', 'scales', 'ggbreak', 'ggforce','gridExtra',
              # Plot interactive network graph
              'visNetwork',
              # Add interactivity to ggplot charts
              'plotly', 'gganimate', 'quantmod', 'crosstalk',
              # Load maps & manipulate geo-spatial data
              'jpeg', 'grid', 'geohashTools', 'sf', 'tmap', 'patchwork', 'raster', 'rgdal', 'mapview')

for (p in packages) {
  if(!require(p, character.only = T)){
    install.packages(p)
  }
  library(p, character.only = T)
}

```

There are 4 csv datasets provided:

- **Credit card data**: transaction of each credit card with timestamp, price and location. 
    + 55 unique credit card numbers
    + 1,490 credit card transactions

```{r cc data}
cc <- read_csv("data/cc_data.csv",
               col_types = cols(timestamp = col_datetime(format = "%m/%d/%Y %H:%M"), 
                                location = col_character(), 
                                price = col_double(), 
                                last4ccnum = col_double()))

cc <- cc %>% arrange(last4ccnum,timestamp) %>%
  mutate(last4ccnum = as.factor(last4ccnum))

glimpse(cc)
```

- **Loyalty card data**: transaction on loyalty card with date, price and location. 
    + 54 unique loyalty card numbers
    + 1,392 loyalty card transactions

```{r loyalty data}
loyalty <- read_csv("data/loyalty_data.csv",
               col_types = cols(timestamp = col_date(format = "%m/%d/%Y"), 
                                location = col_character(), 
                                price = col_double(), 
                                loyaltynum = col_character()))

loyalty <- loyalty %>% arrange(loyaltynum,timestamp) %>%
  mutate(loyaltynum = as.factor(loyaltynum))

glimpse(loyalty)
```

- **GPS data**: timestamped latitude and longitude of each employee vehicle (ID) when in movement, including trucks which `id` is coded in '10X'
    + 40 unique Car IDs (1 to 35, 101, 104 to 107)
    + 685,169 records

```{r gps data}
gps <- read_csv("data/gps.csv",
               col_types = cols(Timestamp = col_datetime(format = "%m/%d/%Y  %H:%M:%S"), 
                                id = col_double(), 
                                lat = col_double(), 
                                long = col_double()))

gps <- gps %>% arrange(id, Timestamp) %>%
  mutate(id = as.factor(id))

glimpse(gps)
```

- **Car assignment data**: full names of employees and ID of cars assigned 
  + 44 known employees
  + 35 cars are given IDs and associated with employees
  + 9 trucks drivers are not assigned with particular vehicle ID

```{r car assignment data}
cars <- read_csv("data/car-assignments.csv",
               col_types = cols(LastName = col_character(), 
                                FirstName = col_character(),
                                CarID = col_double(), 
                                CurrentEmploymentType = col_character(),
                                CurrentEmploymentTitle = col_character()))

cars <- cars %>% arrange(CarID) %>%
  mutate(CarID = as.factor(CarID))

glimpse(cars)
```

Beyond the csv files, a tourist map image and geospatial datasets are provided:


## 5. Data Cleaning and Transformation

### 5.1 Credit and Loyalty Card Transactions

#### 5.1.1 Correct unrecognized character in location name
```{r POI names}
# Match names of locations in credit card and loyalty card transactions to check spelling
cc_loc <- cc %>% distinct(location) %>% arrange(location) 

loyalty_loc <- loyalty %>% distinct(location) %>% arrange(location) 

location <- full_join(cc_loc, loyalty_loc, by = "location") %>% 
            arrange(location)

# Replace the unrecognized characters in Katerina's Cafe location in cc & loyalty data
cc[grep("Katerina", cc$location),2] <- "Katerina's Cafe"
loyalty[grep("Katerina", loyalty$location),2] <- "Katerina's Cafe"
```


#### 5.1.2 Join credit card transactions to loyalty transactions - round 1

Credit card transactions are joined with loyalty card transactions using date, location and spend amount as loyalty card transactions do not have time. 
Any credit card and loyalty card match that only has 1 transaction will be removed to avoid spurious match. From the 1,081 matched transactions, there are 56 unique pairs of credit cards and loyalty cards, which points to multiple relationship between credit card and loyalty cards. As it is hard to visualize the interconnection between credit cards and loyalty cards in tabular form, a network graph will be used to help.

```{r match round 1}
# Create new date field for credit card data
cc <- cc %>% 
  mutate(datestamp = as.Date(timestamp))

# Join datasets on exact date, location and price (round 1)
# 1,807 records with spurious matches and non-matches
transact <- cc %>% 
  full_join(loyalty, by = c("datestamp" = "timestamp", "location", "price"))

# Find the unique pairs of credit card and loyalty card 
# 56 pairs of credit card and loyalty card pairs
pairs <- transact %>% 
  group_by(last4ccnum, loyaltynum) %>%
  count() %>%
  # Remove 6 spurious matches
  filter(n>1) %>%
  drop_na() %>%
  arrange(last4ccnum,loyaltynum) %>%
  ungroup()

# 1,081 matched transactions 
transact_match <- transact %>% 
  drop_na() %>%
  left_join(pairs, by = c("last4ccnum","loyaltynum")) %>% 
  drop_na()%>%
  dplyr::select(-n) 

glimpse(transact_match)
```

 
#### 5.1.2 Visualize cc and loyalty cards relationship using network graph

From the network graph, we observe unexpected relationships between credit card 1286 and loyalty card L3288 and loyalty card L6267 is tied to two credit cards 6691 and 6899. 

```{r prep network graph}
# Prepare cc & loyalty matching data for network graph format
last4ccnum <- transact_match %>%
  distinct(last4ccnum) %>%
  rename(label = last4ccnum) %>%
  mutate(group = 'cc')

loyaltynum <- transact_match %>%
  distinct(loyaltynum) %>%
  rename(label = loyaltynum) %>%
  mutate(group = 'loyalty')

nodes <- full_join(loyaltynum, last4ccnum, by = c("label","group"))

nodes <- nodes %>% rowid_to_column("id") 

per_route <- transact_match %>%  
  group_by(loyaltynum, last4ccnum) %>%
  summarise(weight = n()) %>% 
  ungroup()

edges <- per_route %>% 
  left_join(nodes, by = c("last4ccnum" = "label")) %>% 
  rename(from = id)

edges <- edges %>% 
  left_join(nodes, by = c("loyaltynum" = "label")) %>% 
  rename(to = id)

edges <- dplyr::select(edges, from, to, weight) %>% 
  filter(weight > 1) %>% 
  mutate(width = weight/5 + 1)

# Plot network graph to show cc and loyalty cards relationship
netwkgraph <- visNetwork(nodes, edges, width = "100%", main = list(text = "Relationship Between Credit Card and Loyalty Cards",
 style = "font-family:arial;font-size:20px;")) %>% 
  visOptions(highlightNearest = TRUE,
             nodesIdSelection = list(enabled = TRUE,selected = "1")) %>% 
  visGroups(groupname = "cc", color = "lightblue") %>%    
  visGroups(groupname = "loyalty", color = "orange") %>% 
  visLegend(width = 0.1, position = "right", main = list(text = "Card Types",
 style = "font-family:arial;font-size:15px;"))
netwkgraph
```


#### 5.1.3 Join unmatched credit card transactions to loyalty transactions - round 2

409 loyalty card transactions and 311 credit card transactions were unpaired based on exact date, spend and location. Based on the pairs of credit card and loyalty numbers, we will match the remaining unmatched transactions based on date, location and price range, specifically with loyalty price lower than credit card price. Results show that 219 transactions that have loyalty card price lower than credit card prices. The lower amount registered on the loyalty card usually ranges from $20-80, in increments of $20s. There is no obvious bias to any location or cards. This suggests that there might be a systemic program or error that led to this outcome e.g. cashback program.

```{r match round 2}
# Create config table based on primary matched pairs of loyalty and credit card numbers detected from network graph 
config <- pairs %>%
  filter(!(last4ccnum %in% c('1286') & loyaltynum %in% c('L3288'))) 

# 1,066 transactions with exact match on date, loyalty number, price and location
exact_match <-  cc %>% 
                left_join(config, by = "last4ccnum") %>% 
                inner_join(loyalty, by = c("datestamp" = "timestamp",
                                           "loyaltynum",
                                           "location","price"))

# 409 cc transactions exclude matches from exact price match
cc_unmatched <- cc %>%
  anti_join(transact_match, by = c("datestamp","last4ccnum","location","price")) 

# 311 loyalty transactions exclude matches from exact price match
loyalty_unmatched <- loyalty %>%
  anti_join(transact_match, by = c("timestamp" = "datestamp","loyaltynum","location","price")) 

# Fuzzy match with loyalty price to be lower or equal to cc price returns 496 entries (round 2)
transact_match_2 <- cc_unmatched %>%
  left_join(config, by = "last4ccnum") %>%
  fuzzy_full_join(loyalty_unmatched, 
                  by = c("loyaltynum", "location",
                           "datestamp"="timestamp","price"="price"),
                  match_fun=list(`==`, `==`, `==`, `>=`)) 

# 219 transactions match with price range with price difference in increments of $20.
transact_match_2 <- transact_match_2 %>%
  drop_na() %>%
  mutate(price_diff = price.x - price.y) %>%
  filter(round(price_diff,0) %in% c(20,40,60,80)) 

DT::datatable(transact_match_2, filter = 'top')
```


#### 5.1.4 Join unmatched credit card transactions to loyalty transactions - round 3

The remaining unmatched transactions are matched based on location and price (and not date) to check for any delayed postings of credit card transactions. Results show that 7 Kronos Mart credit card transactions were posted 1 day later than its associated loyalty card transactions.

```{r match round 3}
# 190 cc transactions exclude matches from match round 2 (cashback)
cc_unmatched_2 <- cc_unmatched %>%
  anti_join(transact_match_2, by = c("datestamp", 
                                  "location" = "location.x",
                                  "price" = "price.x",
                                  "last4ccnum")) 

# 92 loyalty transactions exclude matches from match round 2 (cashback)
loyalty_unmatched_2 <- loyalty_unmatched %>%
  anti_join(transact_match_2, by = c("timestamp" = "timestamp.y", 
                                     "loyaltynum" = "loyaltynum.y",
                                     "location" = "location.y",
                                     "price" = "price.y")) 

# Match remaining credit card and loyalty card transactions on location and price (round 3)
transact_match_3 <- full_join(cc_unmatched_2, loyalty_unmatched_2, by = c("location","price"))

# 7 Kronos Mart posts credit card transactions posted 1 day later than loyalty card date
transact_match_3 <- transact_match_3 %>% 
  drop_na() %>%
  mutate(diff_date = datestamp - timestamp.y)

DT::datatable(transact_match_3, filter = 'top')
```


#### 5.1.5 Join unmatched credit card transactions to loyalty transactions - round 4

The 4th attempt to match the remaining unmatched transactions based on location and price range, whereby loyalty card price will be greater than credit card price did not return many matches or a pattern. We will disregard the matches and treat the remainder 183 credit card and 85 loyalty card transactions as situations whereby employees used one of the two cards.

```{r match round 4}
# 183 cc transactions exclude matches from match round 3 (delayed posting)
cc_unmatched_3 <- cc_unmatched_2 %>%
  anti_join(transact_match_3, by = c("datestamp", 
                                  "location",
                                  "price",
                                  "last4ccnum")) 

# 85 loyalty transactions exclude matches from match round 3 (delayed posting)
loyalty_unmatched_3 <- loyalty_unmatched_2 %>%
  anti_join(transact_match_3, by = c("timestamp" = "timestamp.y", 
                                     "loyaltynum" = "loyaltynum",
                                     "location",
                                     "price")) 

# Match with loyalty price to be greater or equal to cc price (round 4)
transact_match_4 <- left_join(cc_unmatched_3,config, by = "last4ccnum") %>%
  fuzzy_full_join(loyalty_unmatched_3, by = c("location",
                                              "price",
                                              "datestamp"="timestamp",
                                              "loyaltynum"),
                  match_fun=list(`==`, 
                              `<=`, 
                              `==`,
                              `==`)) %>%
  drop_na()

DT::datatable(transact_match_4, filter = 'top')

```


#### 5.1.6 Prepare final list of transaction records

The final list of spend transactions yielded 1,575 records, which will be explored for patterns and anomalies.

```{r trans final}

transact_match_clean <- transact_match %>%
  mutate(datestamp_loyalty = datestamp, price_diff = NA, diff_date = NA)

transact_match_2_clean <- transact_match_2 %>%
  dplyr::select(timestamp.x,location.x,price.x,last4ccnum,datestamp,loyaltynum.x,timestamp.y, price_diff) %>%
  rename(timestamp = timestamp.x,
         location = location.x,
         price = price.x,
         loyaltynum = loyaltynum.x,
         datestamp_loyalty = timestamp.y) %>%
  mutate(diff_date = NA)

transact_match_3_clean <- transact_match_3 %>%
  dplyr::select(timestamp.x,location,price,last4ccnum,datestamp,loyaltynum,timestamp.y,diff_date) %>%
  rename(timestamp = timestamp.x, datestamp_loyalty = timestamp.y) %>%
  mutate(price_diff = NA)

loyalty_unmatched_3_clean <- loyalty_unmatched_3 %>%
  rename(datestamp = timestamp) %>%
  mutate(datestamp_loyalty = datestamp)

# Final set of 1,575 credit card and loyalty card transactions
trans_final <- transact_match_clean %>%
  bind_rows(transact_match_2_clean) %>%
  bind_rows(transact_match_3_clean) %>% 
  bind_rows(loyalty_unmatched_3_clean) %>% 
  bind_rows(cc_unmatched_3)

DT::datatable(trans_final, filter = 'top')
```


#### 5.1.7 Understand hourly spend patterns

```{r hourly trans pattern}
# Summarize credit card data by hour
cc_hour <- trans_final %>% 
  drop_na(timestamp) %>%
  mutate(hour_cc = as.numeric(get_hour(timestamp))) %>% 
  group_by(location,hour_cc) %>% 
  count() %>%
  rename("transactions" = n) %>%
  ungroup() %>%
  complete(location, hour_cc = full_seq(hour_cc, period = 1)) %>%
  mutate(hour_cc = as.character.numeric_version(hour_cc)) 

# Plot cc hourly data as heatmap using geom_tile()
heatmap_cc <- ggplot(cc_hour, aes(x= hour_cc, y = location, fill = transactions, 
                                  text = paste('No. of Transactions:', transactions,
                                              '<br>Transaction Hour: ', hour_cc,
                                              '<br>Location: ', location))) +
  theme_minimal() +
  geom_tile(colour="white") +
  labs(x = "Hour", y = "Location") +
  scale_fill_gradient(low = "#E0E1E4", high = "#000000") +
  scale_y_discrete(limits = rev) +
  labs(title = "Hourly Credit Card Transactions",
      x = "Hour", fill = "Transactions") 

# Make heatmap interactive with plotly for exploration
heatmap_cc_gg <- ggplotly(heatmap_cc, tooltip = c("text"))
heatmap_cc_gg
```


#### 5.1.8 Understand daily spend patterns

```{r daily trans patterns}
# Summarize spend data by day
trans_day <- trans_final %>% 
  mutate(date_min = pmin(datestamp, datestamp_loyalty, na.rm=TRUE)) %>%
  group_by(location,date_min) %>% 
  count() %>%
  rename("transactions" = n) %>%
  ungroup() 

break_vec <- trans_day$date_min %>% c(seq(from = min(trans_day$date_min), to = max(trans_day$date_min),by = "day"))

# Plot cc hourly data as heatmap using geom_tile()
heatmap_spend <- ggplot(trans_day, aes(x= date_min, y = location, fill = transactions, 
                                  text = paste('No. of Transactions:', transactions,
                                              '<br>Transaction Date: ', date_min,
                                              '<br>Location: ', location))) +
  theme_minimal() +
  theme(axis.text.x=element_text(angle=45, hjust=1)) +
  geom_tile(colour="white") +
  labs(x = "Date", y = "Location") +
  scale_fill_gradient(low = "#E0E1E4", high = "#000000") +
  scale_y_discrete(limits = rev) +
  scale_x_date(breaks = break_vec, date_label = "%a \n%d %b") +  
  labs(title = "Daily Spend Transactions",
      x = "Date", fill = "Transactions") 

# Make heatmap interactive with plotly for exploration
heatmap_spend_gg <- ggplotly(heatmap_spend, tooltip = c("text"))
heatmap_spend_gg
```

We observed that some locations only have credit card transactions posted in 1 hour within the day e.g. 

- Bean There Done That and Brewed Awakenings, Jack's Magical Beans, Coffee Shack at 12PM
- Daily Dealz at 6AM. 

Another anomaly that we will investigate is the posting from Kronos Mart at 3AM.


#### 5.1.9 Understand spend amount

```{r trans amount}

d <- highlight_key(trans_final)

spend_cc <-  ggplot(d, aes(x = last4ccnum, y = price, color = location), text = paste('Last 4 Credit Card Number', last4ccnum,
                                              '<br>Price: $', price,
                                              '<br>Date: $', datestamp,
                                              '<br>Location: ', location)) +
  geom_point() + 
  theme_minimal() + 
  theme(axis.text.x=element_text(angle=45, hjust=1, size=5)) +
  labs(title = "Spend Amount By Credit Card Number",
      x = "Credit Card Number", y = "Price ($)", color = "Location") 

gg_cc <- highlight(ggplotly(spend_cc), "plotly_selected")

crosstalk::bscols(gg_cc, DT::datatable(d), widths = 12)
```

#### 5.1.10 Zoom in on credit card 9551

```{r daily spend - 9551 patterns}

```







### 5.2 GPS Logs and Map Layers


#### 5.2.1 Set up map layers using the tourist map provided
```{r map layers}
# Import MC2-tourist.tif created using QGIS
ap <- raster("data/Geospatial/MC2-tourist.tif")

# Plotting raster layer using tmap to merge rgb bands
tm_shape(ap) +
tm_rgb(ap, r=1,g=2,b=3,
       alpha = NA,
       saturation = 1,
       interpolate = TRUE,
       max.value = 255)

# Use sf to import Vector GIS data files in ESRI shapefile format
abila_st <- st_read(dsn = "data/Geospatial",
                    layer = "Abila")

```

#### 5.2.2 Create movement path from GPS logs 
```{r gps prep}
# Converting aspatial data into simple feature data frame using st_as_sf
gps_sf <- st_as_sf(gps,
                   coords = c("long","lat"),
                            # EPSG 4326 stands for wgs84 geo. coord sys
                            crs = 4326)

# Group paths by ID and cast as linestring
gps_path <- gps_sf %>%
            group_by(id) %>%
            # Due to peculiar nature of group by, we need to do an action such as summarize
            summarize(m = mean(Timestamp),
                      do_union = FALSE) %>%
            st_cast("LINESTRING") %>%
            ungroup()
# Have to slice the paths by days, or time period to derive meaningful paths
```

#### 5.2.3 Visualize gps paths by Car ID to detect gps anomalies

Results from the maps show that Car ID 28 and 9 have GPS data issues:

- Car ID 28's GPS coordinates are systematically misaligned. Correction can be done by systematically shifting all GPS coordinates using GASTech building as the anchoring position and averaging out the coordinates
- Car ID 9's GPS records are spotty, journeys did not end and start in the same place. Correction can be done by extrapolating journeys. 

```{r plot gps}
# Plot gps paths by car IDs
tmap_mode("view")
tm_shape(ap) +
  tm_rgb(ap, r=1,g=2,b=3,
         alpha = NA,
         saturation = 1,
         interpolate = TRUE,
         max.value = 255) +
tm_shape(gps_path) +
  tm_lines() +
tm_facets(by = "id") 
```

Closer look at GPS of Car 28 and 9:
```{r faulty gps of 28 & 9}
# Zoom in on car 28 & 9
gps_path_selected <- gps_path %>% filter(id %in% c("28","9"))

tmap_mode("view")

tm_shape(ap) +
  tm_rgb(ap, r=1,g=2,b=3,
         alpha = NA,
         saturation = 1,
         interpolate = TRUE,
         max.value = 255) +
tm_shape(gps_path_selected) +
  tm_lines() +
tm_facets(by = "id")
```

#### 5.2.4 Resolve data issues in Car 28 GPS log by taking the average coordinates over a wider time interval

Car ID 28's jittery GPS logs resolved by taking the mean coordinates over a 20 second interval. 
```{r fix gps 28}
# Change GPS timestamp to POSIX format
gps$Timestamp <- as.POSIXct(gps$Timestamp, format = "%m/%d/%Y  %H:%M:%S")

# Group timestamp in intervals of 20 seconds and select the mean coordinates to smoothen out the path
gps_28 <- gps %>%
  filter(id == "28") %>%
  mutate(Timestamp = round_date(Timestamp, "20 sec")) %>%
  mutate(lat = round(lat - 00.004,3) ) %>% 
  mutate(long = round(long + 00.004,3) ) %>%
  group_by(Timestamp) %>%
  summarise(lat = mean(lat, na.rm = FALSE), long = mean(long, na.rm = FALSE)) %>% 
  mutate(id = "28") %>%
  ungroup() 

# Visualize car 28 GPS movement
gps_sf_28 <- st_as_sf(gps_28,
                   coords = c("long","lat"),
                            # EPSG 4326 which stands for wgs84 geo. coord sys
                            crs = 4326)

gps_path_28 <- gps_sf_28 %>%
  group_by(id) %>%
  # Due to peculiar nature of group by, we need to do an action such as summarize
  summarize(m = mean(Timestamp),
            do_union = FALSE) %>%
  st_cast("LINESTRING") %>%
  ungroup()

tmap_mode("view")

tm_shape(ap) +
  tm_rgb(ap, r=1,g=2,b=3,
         alpha = NA,
         saturation = 1,
         interpolate = TRUE,
         max.value = 255) +
tm_shape(gps_path_28) +
  tm_lines()

```

#### 5.2.5 Resolve data issues in Car 9 GPS log

Car ID 9 missing GPS records seems to occur randomly. Unfortunately, we are unable to extrapolate the missing GPS records with safe assumptions. We would take note to register events of Car ID 9 by taking the location when the car stopped or start, whichever that makes more sense (i.e. closer to a meaningful location). 
```{r fix gps 9}
# Understand car ID 9's GPS pattern
gps_9 <- gps %>% 
  mutate(datestamp = as_date(Timestamp)) %>%
  mutate(time = as.POSIXct(Timestamp,format = "%H:%M:%S")) %>%
  filter(id == "9", datestamp == "2014-01-07") 

# Plot GPS pattern in scatterplot chart
gps_9_line <- gps_9 %>%
  ggplot(aes(x = time, y = lat)) +
  geom_point() +
  geom_line(color = "red") +
  facet_grid(row = vars(datestamp), scales = "free") +
  labs(title = "Latitude Coordinates of Car ID 9",
                   x = "Time", y = "Latitude")  

ggplotly(gps_9_line)  
```

#### 5.2.6 Detect events from GPS log

Stops between GPS logs provides clues to where the person was heading to. This study will define all stops between GPS log records that spans above 5 minutes as events to be investigated. Each event will have a time span, a location, and a person tied to it. As a rule of thumb, homes of employee will be first identified as locations with pauses between GPS logs exceeding 5 hours for further investigation. 

```{r event prep, include = TRUE, warning = FALSE}
# Replace gps records of Car ID 28
gps_28 <- gps_28 %>%
  dplyr::select(Timestamp, id, lat, long)

gps_clean <- gps %>%
  filter(!id == "28") %>%
  rbind(gps_28) 

# Make car assignment table neat
cars_neat <- cars %>% 
  unite("name", FirstName, LastName, sep = " ") %>%
  rename(department = CurrentEmploymentType , title = CurrentEmploymentTitle)

# Merge car assignments to gps data & sort table by id & timestamp
gps_name <- gps_clean %>%
  left_join(cars_neat, by = c("id" = "CarID")) %>%
  arrange(id, Timestamp) %>%
  mutate(Timestamp = as.POSIXct(Timestamp, format = "%m/%d/%Y  %H:%M:%S"))

# Find difference between timestamp and id
gps_name_temp <- dplyr::select(gps_name,Timestamp)

time_diff <- tail(gps_name_temp,-1) - head(gps_name_temp,-1)

# Join & Tag Start and Stop for Intervals > 5 min
gps_name_diff <- bind_cols(time_diff$Timestamp,tail(gps_name,-1)) %>%
  rename(diff=...1) %>%
  group_by(id) %>%
  
  # If interval btw timestamps is more than 5 mins, tag as an event
  mutate(event_stop = ifelse(diff/60 > 5,'stop',
                          ifelse(diff/60 < 0,'stop',NA))) %>%
  mutate(event_start = lead(event_stop)) %>% 
  
  # If interval between timestamps is more than 5 hours, tag as home to help identify home locations later
  mutate(home = ifelse(diff/60/60 > 5,'home',NA)) %>%
  mutate(diff = seconds_to_period(diff)) %>%
  mutate(event_start = ifelse(event_start == 'stop','start',NA)) %>%
  
  # Encode lat long into geohash level 7    
  mutate(geohash = gh_encode(latitude = lat, 
                               longitude = long, 
                               precision = 7L)) %>%
  ungroup() 
```

#### 5.2.7 Identify events of Car ID 9 and 28
There is a need to consciously merge the start and stop locations of Car ID 9 as there are places of the interest that are registered by starts (e.g. Bean There Done That) and places registered by stops (e.g. U Pump). We will do the event patching if events identified cannot be matched to a location. 

```{r car 9 events}
# Inspect events of Car ID 9
events_9 <- gps_name_diff %>% 
          filter(id == "9", event_stop == "stop" | event_start == "start") %>%
          unite(event, c("event_start", "event_stop")) %>%
          dplyr::select(Timestamp, lat, long, event)

# view the locations over the jpeg map
map <- ggplot(events_9, aes(long, lat)) + 
    
    theme(panel.border = element_blank(),
    panel.grid.major = element_blank(), 
    panel.grid.minor = element_blank(), 
    panel.background = element_blank(),
    axis.line = element_line(colour = "grey")) +

  # Add points
    geom_point(aes(color=event), size = 3, alpha = 0.8) +
  
  # Add label
    labs(title = "Start and Stop Coordinates of Car ID 9 Events",
                   x = "Longitude", y = "Latitude", color = "Type of Events")

# Use ggplotly to make it interactive so we can pick the right geohash visually
pp <- ggplotly(map)
layout(pp, images = list(
  list(
    source = "https://raw.githubusercontent.com/kpokp/ISSS608_blog/main/_posts/2021-06-10-individual-assignment/images/MC2-tourist.jpg",
    opacity = 0.7,
    layer = "below",
    xref = "x",
    yref = "y",
    sizing="stretch",
    x= 24.8244,
    y= 36.0952,
    sizex= 0.0852,
    sizey= 0.05
  )
))

```

```{r car 28 events}
# Inspect events of Car ID 9
events_28 <- gps_name_diff %>% 
          filter(id == "28", event_stop == "stop" | event_start == "start") %>%
          unite(event, c("event_start", "event_stop")) %>%
          dplyr::select(Timestamp, lat, long, event)

# view the locations over the jpeg map
map <- ggplot(events_28, aes(long, lat)) + 
    
    theme(panel.border = element_blank(),
    panel.grid.major = element_blank(), 
    panel.grid.minor = element_blank(), 
    panel.background = element_blank(),
    axis.line = element_line(colour = "grey")) +

  # Add points
    geom_point(aes(color=event), size = 3, alpha = 0.8) +
  
  # Add label
    labs(title = "Start and Stop Coordinates of Car ID 28 Events",
                   x = "Longitude", y = "Latitude", color = "Type of Events")

# Use ggplotly to make it interactive so we can pick the right geohash visually
pp <- ggplotly(map)
layout(pp, images = list(
  list(
    source = "https://raw.githubusercontent.com/kpokp/ISSS608_blog/main/_posts/2021-06-10-individual-assignment/images/MC2-tourist.jpg",
    opacity = 0.7,
    layer = "below",
    xref = "x",
    yref = "y",
    sizing="stretch",
    x= 24.8244,
    y= 36.0952,
    sizex= 0.0852,
    sizey= 0.05
  )
))

```

#### 5.2.8 Capture all events 

3,067 events with locations that would be later identified as home, places of interest, or unknown/suspicious locations.

```{r events}
# 3,067 events that span more than 5 minutes
events <- gps_name_diff %>%
  filter(event_stop == "stop") %>%
  mutate(event_start_time = Timestamp - diff) %>%
  rename(event_stop_time = Timestamp) %>% 
  mutate(event_date = as_date(event_start_time)) %>% 
  dplyr::select(event_stop, home,event_date, event_start_time, event_stop_time,
                        diff, lat, long, geohash, id, name, department, title) %>%
  filter(diff > 0)

glimpse(events)
```

#### 5.2.8 Identify home locations of employees

The interactive map shows that some of the employees stay together e.g. employees with Car ID 30, 22 and 23. 

Furthermore, due to GPS issues, employee with Car ID 28 has home locations spread around a concentrated area, while employee with Car ID 9 spreads out over a wider area. 

We also picked up GASTech as a location especially for truck drivers, who park their vehicles at the company at the end of their work day.

```{r home events}
# Identify homes or place of accommodation of employees - 500 events
home <- events %>% 
  filter(home == "home") 

home_freq <- home %>%
  dplyr::select(id, lat, long, geohash) %>%
  mutate(lat = round(lat,3), long = round(long,3)) %>%              
  group_by(id, lat, long, geohash) %>%
  summarise(n = n()) %>%
  ungroup()

# view the home locations over the jpeg map
map <- ggplot(home_freq, aes(long, lat, text = paste("geohash:", geohash))) + 
    
    theme(panel.border = element_blank(),
    panel.grid.major = element_blank(), 
    panel.grid.minor = element_blank(), 
    panel.background = element_blank(),
    axis.line = element_line(colour = "grey")) +

  # Add points
    geom_point(aes(color=id, size = n), alpha = 0.8) +
  
  # Add label
    labs(title = "Likely Home Locations by Car IDs",
                   x = "Longitude", y = "Latitude", color = "Car ID", size = "Size of Bubbles - Frequency")

# Use ggplotly to make it interactive so we can pick the right geohash visually
pp <- ggplotly(map)
layout(pp, images = list(
  list(
    source = "https://raw.githubusercontent.com/kpokp/ISSS608_blog/main/_posts/2021-06-10-individual-assignment/images/MC2-tourist.jpg",
    opacity = 0.7,
    layer = "below",
    xref = "x",
    yref = "y",
    sizing="stretch",
    x= 24.8244,
    y= 36.0952,
    sizex= 0.0852,
    sizey= 0.05
  )
))

```


#### 5.2.9 Identify likely geohash of employees' homes locations 
```{r home config}
# Home location config table
location_labels_home <- home_freq %>% 
  group_by(id) %>%
  arrange(desc(n)) %>%
  slice_head() %>% 
  ungroup() %>%
  dplyr::select(geohash, id) %>%
  group_by(geohash) %>% 
  mutate(location = paste0(id, collapse = "/")) %>%
  arrange(geohash) %>%
  slice_head() %>% 
  mutate(label = "Home") %>%
  mutate(location = paste(label,location, sep = " ")) %>%
  dplyr::select(!c(id,label)) %>%
  ungroup()
  
location_labels_home <- home %>% 
  left_join(location_labels_home, by = "geohash") %>%
  drop_na() %>% 
  group_by(location) %>%
  arrange(location) %>%
  slice_head() %>%
  ungroup() %>% 
  mutate(category = "Home") %>%
  dplyr::select("lat","long","geohash","location","category") %>%
  mutate(location = ifelse(geohash == "sw3tn4k", "GASTech", 
                                  ifelse(geohash == "sw3tnw2", "Chostus Hotel",location))) %>%
  mutate(category = ifelse(location == "GASTech", "Company", 
                                  ifelse(location == "Chostus Hotel", "Leisure","Home"))) 

# view the POIs over the jpeg map
map <- ggplot(location_labels_home, aes(long, lat, text = paste('Geohash:', geohash,
                         '<br>Latitude: ', lat,
                         '<br>Longitude: ', long,
                         '<br>Location: ', location))) + 
    
    theme(panel.border = element_blank(),
    panel.grid.major = element_blank(), 
    panel.grid.minor = element_blank(), 
    panel.background = element_blank(),
    axis.line = element_line(colour = "grey")) +

  # Add points
    geom_point(aes(color=location), alpha = 0.8, size =4) +
  
  # Add label
    labs(title = "Accommodation Locations by Car IDs",
                   x = "Longitude", y = "Latitude", color = "Accommodation")

# Use ggplotly to make it interactive so we can pick the right geohash visually
pp <- ggplotly(map, tooltip = "text")
layout(pp, images = list(
  list(
    source = "https://raw.githubusercontent.com/kpokp/ISSS608_blog/main/_posts/2021-06-10-individual-assignment/images/MC2-tourist.jpg",
    opacity = 0.7,
    layer = "below",
    xref = "x",
    yref = "y",
    sizing="stretch",
    x= 24.8244,
    y= 36.0952,
    sizex= 0.0852,
    sizey= 0.05
  )
))
```

#### 5.2.10 Identify relationships of employees who share homes  

From the parallel coordinate plot, we see that 8 of the 15 employees who stay in shared homes are from the Security department. 2 of the homes - Home 13/15/16/21 and Home 22/23/30 are occupied only by employees of the Security department. Home 13/15/16/21 occupants take up Perimeter and Site Control duties while Home 22/23/30 occupants are from the Badging Office and one of them is a Security Group Manager. The other 2 homes make up of employees from varied departments such as Engineering, Facilities, IT and Security IT. 

```{r relationships between shared homes}
# Prepare datatset for parallel coordinates plot
cars_home <- cars_neat %>%
  filter(CarID %in% c("22","23","30","6","25","29","17","24","33","13","15","16","21","14","18")) %>%
  mutate(location = ifelse(CarID %in% c("22","23","30"),"Home 22/23/30",
                           ifelse(CarID %in% c("6","25","29"),"Home 6/25/29",
                           ifelse(CarID %in% c("14","18"),"Home 14/18",
                           ifelse(CarID %in% c("17","24","33"),"Home 17/24/33",
                           ifelse(CarID %in% c("13","15","16","21"),"Home 13/15/16/21",NA)))))) %>%
  rename(a.CarID = CarID, b.name = name, c.location = location, d.department = department, e.title = title) %>%
  group_by(b.name, a.CarID, d.department, e.title, c.location) %>%
  summarise(freq = n()) %>%
  gather_set_data(1:5) %>%
  ungroup()

# Plot parallel coordinate plot
home_para_cord <- ggplot(cars_home, aes(x, id = id, split = y, value = freq)) +
  geom_parallel_sets(aes(fill = c.location), alpha = 0.3, axis.width = 0.1) +
  geom_parallel_sets_axes(axis.width = 0.1, fill = "white", colour = "lightgrey") +
  geom_parallel_sets_labels(colour ="black",angle = 360,size = 3) +
  theme_minimal() +
  theme(panel.grid.major = element_blank(),
        axis.text.y = element_blank(),
        axis.line = element_blank()) +
 # Add label
  labs(title = "Characteristics of Employees Sharing Homes",
       x = "Characteristics", fill = "Home Locations") 
  #facet_grid(~location)

home_para_cord
  
```

#### 5.2.11 Identify anomalies from home stays

From the interactive timeline, we observe some activities that stand out for further investigation: 
- ID 1 was not home for 2 nights (6 Jan and 8 Jan)
- ID 16 was staying overnight in Aliba starting 7 Jan 
- ID 21 visited 2 homes from 10 Jan (Home 21 and 35) and visited an unknown location subsequently. Home location was also not detected on night of 14 Jan.
- ID 22 and 28 visited unknown locations in the afternoon of 18 Jan
- ID 24 was not home on 9 Jan
- ID 25 visited an unknown location on 18 Jan that lasted for almost 24 hours
- ID 31 was only in Aliba for 4day-3nights (16 to 19 Jan) and stayed at the Chostus hotel
- Selected trucks were only parked at GASTech building from 6 Jan to 17 Jan, and they are parked at the GASTech building

```{r home matched}
# Match home events to home config table based on 150m radius and no duplicate was introduced - 494 events matched, 6 unmatched
# Unknown locations will be investigated
home_matched <- home %>%
  geo_left_join(location_labels_home, by = c("lat", "long"), max_dist = 0.15, distance_col = "distance", unit = c("km")) %>% 
  # Assume employee with car ID 9 to be home
  mutate(location = ifelse(id == 9, "Home 9",location)) %>%
  mutate(category = ifelse(id == 9, "Home",location)) 
  
# Plot timeline data as gantt chart
home_timeline <- ggplot(home_matched,
            aes(text = paste('id:', id,
                         '<br>Event Start Time: ', as.POSIXct(event_start_time, format = "%m/%d/%Y  %H:%M:%S"),
                         '<br>Event Stop Time: ', as.POSIXct(event_stop_time, format = "%m/%d/%Y  %H:%M:%S"),
                         '<br>Location: ', location))) +
  geom_linerange(mapping=aes(y=id, xmin=event_start_time, xmax=event_stop_time, color=location),size=I(1)) +
  scale_y_discrete(limits = rev) +
  scale_x_datetime(date_breaks="1 day") + 
  theme_minimal() +
  theme(axis.text.x=element_text(angle=45, hjust=1)) +
  # Add label
  labs(title = "Home Locations by Car IDs",
       x = "Date", y = "Car ID", color = "Home Locations") 
  # Can add highlight when do not need tooltip
  # gghighlight(is.na(location))


# Make gantt chart interactive
ggplotly(home_timeline, tooltip = c("text"))
```



### 5.3 Combine Spending Transactions with Events Detected from GPS

#### 5.3.1 Find out the lat long of credit card events based on time frame
- If the transaction time is within the time span of an event, it is assigned to that event

```{r event and transactions}
# Select events that are not staying at home by excluding home - 2,567 events
POI <- events %>%
  filter(is.na(home)) 

# Match POI events with home locations to narrow list of locations to match
POI_matched_home <- POI %>%
  left_join(location_labels_home, by = "geohash")

# Add the unknown locations from home events match to the rest of the events - 7 events
unknown_home_location <- home_matched %>% filter(is.na(location)) %>%
  dplyr::select(event_stop,home,event_date, event_start_time,event_stop_time,diff,lat.x,long.x,geohash.x,id,name,department,title) %>%
  rename(geohash = geohash.x)

# List of events to be matched
POI_2 <- POI_matched_home %>%
  filter(is.na(location)) %>%
  bind_rows(unknown_home_location)

# Select distinct timestamp with each location from transactions
#trans_final
trans_time <- cc %>%
  dplyr::select(location, timestamp) %>%
  distinct(location, timestamp) 

# Remove locations that coincided with Home locations and add the unknown events from Home matching              
# Perform fuzzy match of rest of the events and transactions by time windows
POI_matched <- POI_2 %>%
  fuzzy_left_join(trans_time,
                  by = c("event_start_time" = "timestamp", "event_stop_time" = "timestamp"),
                  match_fun = list(`<`,`>`)) 

# Prepare the matched data for visualization - 996 records 
# We would not pick the top matches as some locations are less visited.
POI_matched_2 <- POI_matched %>% 
  dplyr::select(!home) %>% 
  drop_na(location.y) %>% 
  # Round the lat long to reduce effect of spotty gps
  mutate(lat = round(lat.x,3), long = round(long.x,3)) %>% 
  # Remove Car ID 9 due to spotty GPS
  filter(!id == "9") %>%
  dplyr::select(location.y, lat, long, geohash) %>% 
  group_by(location.y, lat, long, geohash) %>% 
  summarise(n = n()) %>%
  ungroup() %>% 
  arrange(location.y, desc(n)) %>% 
  group_by(location.y) %>%
  ungroup() 

glimpse(POI_matched_2)

```


#### 5.3.2 Plot the events that coincide with transactions to determine the lat long of POIs
The interactive map shows locations where credit card and gps locations coincide. We notice that the highest frequency matches need not coincide with the POI placement on the map. As such, we visually pick out the geohash of each POI based on the the description on the map, coupled with the highest occurrence. 

```{r POI coordinates, include = TRUE, warning = FALSE}
# view the POIs over the jpeg map
map <- ggplot(POI_matched_2, aes(long, lat, text = paste("geohash:", geohash))) + 
    
    theme(panel.border = element_blank(),
    panel.grid.major = element_blank(), 
    panel.grid.minor = element_blank(), 
    panel.background = element_blank(),
    axis.line = element_line(colour = "grey")) +

  # Add points
  geom_point(aes(color=location.y, size = n), alpha = 0.8) +
  
  # Add label
  labs(title = "Likely POI Locations",
                   x = "Longitude", y = "Latitude", color = "Location", size = "Size of Bubbles - Frequency")
 
# Use ggplotly to make it interactive so we can pick the right geohash visually
pp <- ggplotly(map)
layout(pp, images = list(
  list(
    source = "https://raw.githubusercontent.com/kpokp/ISSS608_blog/main/_posts/2021-06-10-individual-assignment/images/MC2-tourist.jpg",
    opacity = 0.5,
    layer = "below",
    xref = "x",
    yref = "y",
    sizing="stretch",
    x= 24.8244,
    y= 36.0952,
    sizex= 0.0852,
    sizey= 0.05
  )
))
```


Most likely geohashes identified per location:

- Abila Airport: sw3thfv
- Abila Scrapyard: sw3tjrh
- Abila Zacharo: sw3tjmp
- Ahaggo Museum: sw3tnpd
- Albert's Fine Clothing: sw3tjx7
- Bean There Done That: sw3tm2r
- Brew've Been Served: sw3tnek 
- Carlyle Chemical Inc.: sw3tnhm
- Coffee Cameleon: sw3tn7s
- Desafio Golf Course: sw3tm9y
- Frank's Fuel: sw3tjq8
- Frydos Autosupply n' More: sw3tnes 
- Gelatogalore: sw3tjsm
- General Grocer: sw3tjse
- Guy's Gyros: sw3tnev 
- Hallowed Grounds: sw3tnm1
- Hippokampos: sw3tjxs
- Jack's Magical Beans: sw3tnjb
- Kalami Kafenion: sw3tjt8
- Katerina's Cafe: sw3tnee 
- Kronos Mart: sw3tjmx
- Kronos Pipe and Irrigation: sw3tjj6
- Maximum Iron and Steel: sw3tjm2
- Nationwide Refinery: sw3tnk1 
- Octavio's Office Supplies: sw3tjsk
- Ouzeri Elian: sw3tjgn
- Roberts and Sons: sw3tjt0
- Shoppers' Delight: sw3tjgh
- Stewart and Sons Fabrication: sw3tng3
- U-Pump: sw3tjvy
- Kronos Capitol: sw3tj7n

#### 5.3.3 Create a config table that holds the likely geohashes of each POI 
```{r POI config}
# Create a config table of the POI with its most likely geohash
POI_geohash <- tibble(location = c("Abila Airport","Abila Scrapyard","Abila Zacharo","Ahaggo Museum","Albert's Fine Clothing",
                                   "Bean There Done That","Brew've Been Served","Carlyle Chemical Inc.","Coffee Cameleon",
                                   "Desafio Golf Course","Frank's Fuel","Frydos Autosupply n' More",
                                   "Gelatogalore","General Grocer","Guy's Gyros","Hallowed Grounds","Hippokampos",
                                   "Jack's Magical Beans","Kalami Kafenion","Katerina's Cafe","Kronos Mart","Kronos Pipe and Irrigation",
                                   "Maximum Iron and Steel","Nationwide Refinery","Octavio's Office Supplies","Ouzeri Elian","Roberts and Sons",
                                   "Shoppers' Delight","Stewart and Sons Fabrication","U-Pump","Kronos Capitol"),
                      geohash_likely = c("sw3thfv","sw3tjrh","sw3tjmp","sw3tnpd","sw3tjx7",
                                        "sw3tm2r","sw3tnek","sw3tnhm","sw3tn7s",
                                        "sw3tm9y","sw3tjq8","sw3tnes",
                                        "sw3tjsm","sw3tjse","sw3tnev","sw3tnm1","sw3tjxs",
                                        "sw3tnjb","sw3tjt8","sw3tnee","sw3tjmx","sw3tjj6",
                                        "sw3tjm2","sw3tnk1","sw3tjsk","sw3tjgn","sw3tjt0",
                                        "sw3tjgh","sw3tng3","sw3tjvy","sw3tj7n")) 

POI_geohash_2 <- POI_matched_2 %>%
  left_join(POI_geohash, by = c("geohash" = "geohash_likely")) %>% 
  group_by(location) %>%
  arrange(desc(n)) %>%
  slice_head() %>%
  ungroup() %>%
  dplyr::select(lat, long, geohash, location) %>%
  drop_na() %>%
  
  # as the cc timing did not match gps due to cc posting issues or non-spend locations
  add_row(lat = 36.073, long = 24.864, geohash = "sw3tjycr", location = "Brewed Awakenings")
  # add_row(lat = 36.066, long = 24.850, geohash = "sw3tjmx", location = "Kronos Mart")

# Assign categories to add semantics to points-of-interest (POI)
food <- c("Abila Zacharo","Bean There Done That","Brew've Been Served",
          "Brewed Awakenings","Coffee Cameleon","Coffee Shack","Gelatogalore",
          "Guy's Gyros", "Hallowed Grounds","Hippokampos","Jack's Magical Beans", 
          "Kalami Kafenion", "Katerina's Cafe","Ouzeri Elian")
retail <- c("Albert's Fine Clothing","Daily Dealz","General Grocer","Kronos Mart",
            "Roberts and Sons", "Shoppers' Delight")
gas <- c("Frank's Fuel","U-Pump")
leisure <- c("Abila Airport","Ahaggo Museum","Chostus Hotel","Desafio Golf Course","Kronos Capitol")
company <- c("Abila Scrapyard","Carlyle Chemical Inc.","Frydos Autosupply n' More",
             "Kronos Pipe and Irrigation","Maximum Iron and Steel","GASTech",
             "Nationwide Refinery","Octavio's Office Supplies","Stewart and Sons Fabrication")

# Add categories to config table
location_labels_POI <- POI_geohash_2 %>% mutate(category = 
                    ifelse(location %in% food, "Food",
                    ifelse(location %in% retail, "Retail",
                    ifelse(location %in% gas, "Gas",
                    ifelse(location %in% leisure, "Leisure",
                    ifelse(location %in% company, "Company", NA)))))) 

# Check that POI positions are correctly tagged on the map
map <- ggplot(location_labels_POI, aes(long, lat, text = paste('Category:', category,
                                                        '<br>Location: ', location,
                                                        '<br>Geohash: ', geohash,
                                                        '<br>Lat: ', lat,
                                                        '<br>Long: ', long))) + 
    
    theme(panel.border = element_blank(),
    panel.grid.major = element_blank(), 
    panel.grid.minor = element_blank(), 
    panel.background = element_blank(),
    axis.line = element_line(colour = "grey")) +

  # Add points
  geom_point(aes(color= category), alpha = 0.8, size = 3) +
  
  # Add label
  labs(title = "POI Locations", x = "Longitude", y = "Latitude", color = "Category")


# Use ggplotly to make it interactive so we can pick the right geohash visually
pp <- ggplotly(map, tooltip = c("text"))

layout(pp, images = list(
  list(
    source = "https://raw.githubusercontent.com/kpokp/ISSS608_blog/main/_posts/2021-06-10-individual-assignment/images/MC2-tourist.jpg",
    opacity = 0.3,
    layer = "below",
    xref = "x",
    yref = "y",
    sizing="stretch",
    x= 24.8244,
    y= 36.0952,
    sizex= 0.0852,
    sizey= 0.05
  )
))

```

#### 5.3.4 Match the rest of the 2,567 events + 7 unknown events with POIs and home locations
```{r POI matched}
# Create a config table with both POI and home locations
location_labels <- location_labels_POI %>% rbind(location_labels_home) 

# Match rest of events to POI config table based on 120m radius and pick the location that is closest to POI using geo-join
# Unknown locations will be investigated
POI_geo_matched <-  POI_2 %>%
  rename(lat = lat.x, long = long.x) %>%
  dplyr::select(-c(lat.y, long.y, location, category)) %>% 
  geo_left_join(location_labels, by = c("lat", "long"), max_dist = 0.12, distance_col = "distance", unit = c("km")) %>% 
  group_by(id, event_stop_time) %>%
  arrange(desc(distance)) %>%
  slice_head()

```

#### 5.3.5 Identify unknown events on the map
```{r unknown events}
# Isolate the unknown events
unknown_location <- POI_geo_matched %>% 
  filter(is.na(location)) %>%
  dplyr::select(event_stop,home,event_start_time,event_stop_time,diff,lat.x,long.x,geohash.x,id,name) %>%
  rename(lat = lat.x, long = long.x, geohash = geohash.x)

map <- ggplot(unknown_location, aes(long, lat, text = paste("geohash:", geohash))) + 
    
    theme(panel.border = element_blank(),
    panel.grid.major = element_blank(), 
    panel.grid.minor = element_blank(), 
    panel.background = element_blank(),
    axis.line = element_line(colour = "grey")) +

  # Add points
  geom_point(aes(color=id), alpha = 0.8, size = 3) +
  
  # Add label
  labs(title = "Unknown Locations by Car IDs", color = "Car ID")

# Use ggplotly to make it interactive so we can pick the right geohash visually
pp <- ggplotly(map)

layout(pp, images = list(
  list(
    source = "https://raw.githubusercontent.com/kpokp/ISSS608_blog/main/_posts/2021-06-10-individual-assignment/images/MC2-tourist.jpg",
    opacity = 0.3,
    layer = "below",
    xref = "x",
    yref = "y",
    sizing="stretch",
    x= 24.8244,
    y= 36.0952,
    sizex= 0.0852,
    sizey= 0.05
  )
))
```

We observe some patterns from the map above:
- Car ID 9's spotty gps cause events to be identified when there is none. We will ignore the events until we fix the journeys of Car ID 9 with its event start locations.
- Excluding events from 9, there are 5 locations that cannot be explained. They are visited by Car ID 13, 15, 21 and 24

#### 5.3.6 Do a match of all events against the 3 catgeories of locations - home, POI and unknown
```{r events match config}
# Create a config for suspicious locations tagged as unknown category
location_labels_unknown <- unknown_location %>% 
  filter(!id %in% c("9")) %>%
  dplyr::select(lat,long,geohash,id) %>%
  mutate(lat = round(lat,2),long=round(long,2)) %>%
  group_by(lat,long,id) %>%
  slice_head() %>%
  ungroup() %>% 
  group_by(lat,long) %>% 
  mutate(location = paste0(id, collapse = "/")) %>%
  arrange(geohash) %>%
  slice_head() %>% 
  dplyr::select(!c(id,event_stop_time)) %>%
  ungroup() %>%
  mutate(label = "Unknown", category = "Unknown") %>%
  mutate(location = paste(label,location, sep = " ")) %>%
  left_join(unknown_location, by = "geohash") %>%
  dplyr::select(lat.y, long.y, geohash, location, category) %>%
  group_by(geohash,location) %>%
  slice_head() %>%
  ungroup() %>%
  rename(lat = lat.y, long = long.y) 

# Add suspicious locations to the location labels config table
location_labels_all <- location_labels %>% bind_rows(location_labels_unknown)

# Clean home events
home_matched <- filter(home_matched, !is.na(location)) %>%
  dplyr::select(event_stop,home,event_date,event_start_time,event_stop_time,diff,lat.x,long.x,geohash.x,id,name,department,title,location) %>%
  mutate(category = "Home") %>%
  rename(lat=lat.x,long=long.x,geohash=geohash.x) 
# Clean home events
POI_matched_home <- filter(POI_matched_home, !is.na(location)) %>%
  dplyr::select(event_stop,home,event_date,event_start_time,event_stop_time,diff,lat.x,long.x,geohash,id,name,department,title,location,category) %>%
  rename(lat=lat.x,long=long.x)    

# Match 1,429 unmatched POI events that is closest distance to POI to avoid duplicates, 12 unmatched events from Car 9
POI_3 <- POI_2 %>%
  dplyr::select(event_stop,home,event_date,event_start_time,event_stop_time,diff,lat.x,long.x,geohash,id,name,department,title) %>%
  geo_left_join(location_labels_all, by = c("lat.x" = "lat", "long.x" = "long"), max_dist = 0.12, distance_col = "distance", unit = c("km")) %>%
  dplyr::select(event_stop,home,event_date,event_start_time,event_stop_time,diff,lat.x,long.x,geohash.x,id,name,department,title,location,category, distance) %>%
  rename(lat=lat.x,long=long.x,geohash=geohash.x) %>%
  group_by(event_start_time, id) %>%
  arrange(distance) %>%
  slice_head() %>% 
  ungroup()

# Match 12 starting POI events of Car9 that is closest distance to POI to avoid duplicates  
POI_Car9 <- gps_name_diff %>%
  filter((event_start == "start"|event_stop == "stop"), id == "9") %>% 
  left_join(filter(POI_3,is.na(location)), by = c("Timestamp"="event_start_time")) %>% 
  filter(!is.na(geohash.y)) %>%
  dplyr::select(event_stop.y,home.y,event_date, Timestamp, event_stop_time,
                        diff.y, lat.x, long.x, geohash.x, id.y, name.y, department.y, title.y) %>%
  rename(event_stop=event_stop.y,home=home.y,event_start_time=Timestamp,
                        diff=diff.y, lat=lat.x, long=long.x, geohash=geohash.x, id=id.y, name=name.y, department=department.y, title=title.y) %>%
  geo_left_join(location_labels_all, by = c("lat","long"), max_dist = 0.25, distance_col = "distance", unit = c("km")) %>% 
  group_by(event_start_time, id) %>%
  arrange(distance) %>%
  slice_head() %>% 
  ungroup() %>%
  dplyr::select(-c(lat.y,long.y,geohash.y)) %>%
  rename(lat=lat.x, long=long.x,geohash=geohash.x,)

# Match all 3,067 events 
events_matched <- filter(POI_3,!is.na(location)) %>%
  # Add 12 events from Car ID 9  
  bind_rows(POI_Car9) %>%
  # Add 493 homes events 
  bind_rows(home_matched) %>%
  # Add another 1,145 POI events matched to home locations
  bind_rows(POI_matched_home)

# Map all locations
map <- ggplot(location_labels_all, aes(long, lat, text = paste("geohash:", geohash))) + 
    
    theme(panel.border = element_blank(),
    panel.grid.major = element_blank(), 
    panel.grid.minor = element_blank(), 
    panel.background = element_blank(),
    axis.line = element_line(colour = "grey")) +

  # Add points
  geom_point(aes(color=location), alpha = 0.8, size = 3) +
  
  # Add label
  labs(title = "Locations", color = "Location")

# Use ggplotly to make it interactive so we can pick the right geohash visually
pp <- ggplotly(map)

layout(pp, images = list(
  list(
    source = "https://raw.githubusercontent.com/kpokp/ISSS608_blog/main/_posts/2021-06-10-individual-assignment/images/MC2-tourist.jpg",
    opacity = 0.3,
    layer = "below",
    xref = "x",
    yref = "y",
    sizing="stretch",
    x= 24.8244,
    y= 36.0952,
    sizex= 0.0852,
    sizey= 0.05
  )
))
```

#### 5.3.7 Plot timeline of events to check if all events are captured

```{r events timeline}
# Change data into long form for line chart
events_matched_long <- events_matched %>% 
  mutate(rownum = row_number()) %>%
  dplyr::select(!event_stop) %>%
  pivot_longer(cols = starts_with("event_st"), names_to = "event_start_stop", values_to = "timestamp")

events_matched_long_highlight <- highlight_key(events_matched_long)

# Plot timeline data as gantt chart
events_timeline <- ggplot(events_matched_long_highlight, 
                          aes(text = paste('id:', id,
                                           '<br>Name:', name,
                                           '<br>Department:', department,
                                           '<br>Tite:', title,
                                           '<br>Timestamp:', timestamp,
                                           '<br>Location: ', location,
                                           '<br>Category: ', category))) +
  # Add line range
  geom_line(mapping=aes(y=id,x=timestamp,color=location,group=rownum),size=1) +
  
  # Change theme
  theme_minimal() + 
  
  # Modify scale
  scale_y_discrete(limits = rev) +
  scale_x_datetime(date_breaks="1 day", labels =label_date_short()) + 
  theme(axis.text.x=element_text(hjust=1)) +
    
  # Add label
  labs(title = "Locations by Car IDs", y = "Car ID", x = "Date", color = "Locations") 
  

# Make gantt chart interactive 
bscols(
  widths = c(12, 12), 
  #filter_select("id", "Select ID", events_matched_long_highlight, ~id, multiple = TRUE),
  #filter_select("category", "Select Location Category", events_matched_long_highlight, ~category),
  ggplotly(events_timeline, tooltip = c("text"))
  )
```

#### 5.3.8 Informal relationships between individuals through home visits

We want to understand the informal relationships between employees through the homes they visit. We isolated the homes from the event timeline and picked out the homes which were visited by people besides their usual occupants. 

- Lars (Home 2) from Engineering hosted a home party on Friday night for 15 colleagues from It and Engineering departments
- Executives are being watched by 4 individuals who are likely connected - Loreto (15), Isia (16), Hennie (21) and Minke (24)
- Hennie (21) visits home of Lidslse (14) and Birgitta (18) every day around dinner time and stays over on Wednesday and weekends
- Isande (28) visits home of Adra, Varja and Felix (22/23/30) on weekend nights

```{r home events timeline}
# Change data into long form for line chart
events_matched_long_home <- events_matched %>% 
  filter(str_detect(location,"Home")) %>%
  mutate(rownum = row_number()) %>%
  dplyr::select(!event_stop) %>%
  pivot_longer(cols = starts_with("event_st"), names_to = "event_start_stop", values_to = "timestamp") 

# Plot timeline data as gantt chart
events_timeline_home <- ggplot(events_matched_long_home, 
                          aes(text = paste('id:', id,
                                           '<br>Name:', name,
                                           '<br>Department:', department,
                                           '<br>Tite:', title,
                                           '<br>Timestamp:', timestamp,
                                           '<br>Location: ', location,
                                           '<br>Category: ', category))) +
  # Add line range
  geom_line(mapping=aes(y=id,x=timestamp,color=location,group=rownum),size=1) +
  
  # Change theme
  theme_minimal() + 
  
  # Modify scale
  scale_y_discrete(limits = rev) +
  scale_x_datetime(date_breaks="1 day", labels =label_date_short()) + 
  theme(axis.text.x=element_text(hjust=1)) +
    
  # Add label
  labs(title = "Locations by Car IDs", y = "Car ID", x = "Date", color = "Locations") 
  
  # Can add highlight when do not need tooltip
  # gghighlight(location == "Home 2") 

# Make gantt chart interactive
ggplotly(events_timeline_home, tooltip = c("text"))
```
```{r home events - House 2 timeline}
# Change data into long form for line chart
events_matched_long_home2 <- events_matched %>% 
  filter(location == "Home 2") %>%
  mutate(rownum = row_number()) %>%
  dplyr::select(!event_stop) %>%
  pivot_longer(cols = starts_with("event_st"), names_to = "event_start_stop", values_to = "timestamp") 

# Plot timeline data as gantt chart
events_timeline_home2 <- ggplot(events_matched_long_home2, 
                          aes(text = paste('id:', id,
                                           '<br>Name:', name,
                                           '<br>Department:', department,
                                           '<br>Tite:', title,
                                           '<br>Timestamp:', timestamp,
                                           '<br>Location: ', location,
                                           '<br>Category: ', category))) +
  # Add line range
  geom_line(mapping=aes(y=id,x=timestamp,color=location,group=rownum),size=1) +
  
  # Change theme
  theme_minimal() + 
  
  # Modify scale
  scale_y_discrete(limits = rev) +
  scale_x_datetime(date_breaks="1 day", labels =label_date_short()) + 
  theme(axis.text.x=element_text(hjust=1)) +
    
  # Add label
  labs(title = "Locations by Car IDs - Friday Dinner Party at Lars", y = "Car ID", x = "Date", color = "Locations") 
  
  # Can add highlight when do not need tooltip
  # gghighlight(location == "Home 2") 

# Make gantt chart interactive
ggplotly(events_timeline_home2, tooltip = c("text"))
```

```{r home events - executive timeline}
# Change data into long form for line chart
events_matched_long_exechome <- events_matched %>% 
  filter(location %in% c("Home 10", "Home 32", "Home 35", "Home 4")) %>%
  mutate(rownum = row_number()) %>%
  dplyr::select(!event_stop) %>%
  pivot_longer(cols = starts_with("event_st"), names_to = "event_start_stop", values_to = "timestamp") 

# Plot timeline data as gantt chart
events_timeline_exechome <- ggplot(events_matched_long_exechome, 
                          aes(text = paste('id:', id,
                                           '<br>Name:', name,
                                           '<br>Department:', department,
                                           '<br>Tite:', title,
                                           '<br>Timestamp:', timestamp,
                                           '<br>Location: ', location,
                                           '<br>Category: ', category))) +
  # Add line range
  geom_line(mapping=aes(y=id,x=timestamp,color=location,group=rownum),size=1) +
  
  # Change theme
  theme_minimal() + 
  
  # Modify scale
  scale_y_discrete(limits = rev) +
  scale_x_datetime(date_breaks="1 day", labels =label_date_short()) + 
  theme(axis.text.x=element_text(hjust=1)) +
    
  # Add label
  labs(title = "Locations by Car IDs - Midnight Surveillance of the Executives' Homes", y = "Car ID", x = "Date", color = "Locations") 

# Make gantt chart interactive
ggplotly(events_timeline_exechome, tooltip = c("text"))
```

```{r home events - friends timeline}
# Change data into long form for line chart
events_matched_long_friendhome <- events_matched %>% 
  filter(location %in% c("Home 14/18", "Home 22/23/30")) %>%
  mutate(rownum = row_number()) %>%
  dplyr::select(!event_stop) %>%
  pivot_longer(cols = starts_with("event_st"), names_to = "event_start_stop", values_to = "timestamp") 

# Plot timeline data as gantt chart
events_timeline_friendhome <- ggplot(events_matched_long_friendhome, 
                          aes(text = paste('id:', id,
                                           '<br>Name:', name,
                                           '<br>Department:', department,
                                           '<br>Tite:', title,
                                           '<br>Timestamp:', timestamp,
                                           '<br>Location: ', location,
                                           '<br>Category: ', category))) +
  # Add line range
  geom_line(mapping=aes(y=id,x=timestamp,color=location,group=rownum),size=1) +
  
  # Change theme
  theme_minimal() + 
  
  # Modify scale
  scale_y_discrete(limits = rev) +
  scale_x_datetime(date_breaks="1 day", labels =label_date_short()) + 
  theme(axis.text.x=element_text(hjust=1)) +
    
  # Add label
  labs(title = "Locations by Car IDs - Dating or Close Friends?", y = "Car ID", x = "Date", color = "Locations") 

# Make gantt chart interactive
ggplotly(events_timeline_friendhome, tooltip = c("text"))
```

#### 5.3.9 Informal relationships between individuals through leisure

We want to understand the informal relationships between employees through the leisure locations they visit. We isolated the Leisure events from timeline.

- Executives - Ingrid (4), Ada (10), Willem (35) and Orhan (32) - play golf on Sunday afternoons and Sten (31) joined when he dropped by Aliba
- Elsa (7) and Brand (33) visited Chostus hotel together on 8, 10, 14 and 17 Jan around lunch time
- Kanon (28) and Borrasca (28) hung out at the museum and park together on 18 Jan

```{r leisure events timeline}
# Change data into long form for line chart
events_matched_long_leisure <- events_matched %>% 
  filter(category == "Leisure") %>%
  mutate(rownum = row_number()) %>%
  dplyr::select(!event_stop) %>%
  pivot_longer(cols = starts_with("event_st"), names_to = "event_start_stop", values_to = "timestamp") 

# Plot timeline data as gantt chart
events_timeline_leisure <- ggplot(events_matched_long_leisure, 
                          aes(text = paste('id:', id,
                                           '<br>Name:', name,
                                           '<br>Department:', department,
                                           '<br>Tite:', title,
                                           '<br>Timestamp:', timestamp,
                                           '<br>Location: ', location,
                                           '<br>Category: ', category))) +
  # Add line range
  geom_line(mapping=aes(y=id,x=timestamp,color=location,group=rownum),size=1) +
  
  # Change theme
  theme_minimal() + 
  
  # Modify scale
  scale_y_discrete(limits = rev) +
  scale_x_datetime(date_breaks="1 day", labels =label_date_short()) + 
  theme(axis.text.x=element_text(hjust=1)) +
    
  # Add label
  labs(title = "Locations by Car IDs", y = "Car ID", x = "Date", color = "Locations") 
  
  # Can add highlight when do not need tooltip
  # gghighlight(location == "Home 2") 

# Make gantt chart interactive
ggplotly(events_timeline_leisure, tooltip = c("text"))
```

```{r leisure events - hotel timeline}
# Change data into long form for line chart
events_matched_long_hotel <- events_matched %>% 
  filter(location == "Chostus Hotel") %>%
  mutate(rownum = row_number()) %>%
  dplyr::select(!event_stop) %>%
  pivot_longer(cols = starts_with("event_st"), names_to = "event_start_stop", values_to = "timestamp") 

# Plot timeline data as gantt chart
events_timeline_hotel <- ggplot(events_matched_long_hotel, 
                          aes(text = paste('id:', id,
                                           '<br>Name:', name,
                                           '<br>Department:', department,
                                           '<br>Tite:', title,
                                           '<br>Timestamp:', timestamp,
                                           '<br>Location: ', location,
                                           '<br>Category: ', category))) +
  # Add line range
  geom_line(mapping=aes(y=id,x=timestamp,color=location,group=rownum),size=1) +
  
  # Change theme
  theme_minimal() + 
  
  # Modify scale
  scale_y_discrete(limits = rev) +
  scale_x_datetime(date_breaks="1 day", labels =label_date_short()) + 
  theme(axis.text.x=element_text(hjust=1)) +
    
  # Add label
  labs(title = "Locations by Car IDs - Afternoon Rendezvous at Chostus Hotel", y = "Car ID", x = "Date", color = "Locations") 

# Make gantt chart interactive
ggplotly(events_timeline_hotel, tooltip = c("text"))
```

```{r leisure events - golf timeline}
# Change data into long form for line chart
events_matched_long_golf <- events_matched %>% 
  filter(location == "Desafio Golf Course") %>%
  mutate(rownum = row_number()) %>%
  dplyr::select(!event_stop) %>%
  pivot_longer(cols = starts_with("event_st"), names_to = "event_start_stop", values_to = "timestamp") 

# Plot timeline data as gantt chart
events_timeline_golf <- ggplot(events_matched_long_golf, 
                          aes(text = paste('id:', id,
                                           '<br>Name:', name,
                                           '<br>Department:', department,
                                           '<br>Tite:', title,
                                           '<br>Timestamp:', timestamp,
                                           '<br>Location: ', location,
                                           '<br>Category: ', category))) +
  # Add line range
  geom_line(mapping=aes(y=id,x=timestamp,color=location,group=rownum),size=1) +
  
  # Change theme
  theme_minimal() + 
  
  # Modify scale
  scale_y_discrete(limits = rev) +
  scale_x_datetime(date_breaks="1 day", labels =label_date_short()) + 
  theme(axis.text.x=element_text(hjust=1)) +
    
  # Add label
  labs(title = "Locations by Car IDs - CEO Joins Executives' Weekly Sunday Golf", y = "Car ID", x = "Date", color = "Locations") 

# Make gantt chart interactive
ggplotly(events_timeline_golf, tooltip = c("text"))
```

```{r leisure events - park timeline}
# Change data into long form for line chart
events_matched_long_park <- events_matched %>% 
  filter(location %in% c("Ahaggo Museum","Kronos Capitol")) %>%
  mutate(rownum = row_number()) %>%
  dplyr::select(!event_stop) %>%
  pivot_longer(cols = starts_with("event_st"), names_to = "event_start_stop", values_to = "timestamp") 

# Plot timeline data as gantt chart
events_timeline_park <- ggplot(events_matched_long_park, 
                          aes(text = paste('id:', id,
                                           '<br>Name:', name,
                                           '<br>Department:', department,
                                           '<br>Tite:', title,
                                           '<br>Timestamp:', timestamp,
                                           '<br>Location: ', location,
                                           '<br>Category: ', category))) +
  # Add line range
  geom_line(mapping=aes(y=id,x=timestamp,color=location,group=rownum),size=1) +
  
  # Change theme
  theme_minimal() + 
  
  # Modify scale
  scale_y_discrete(limits = rev) +
  scale_x_datetime(date_breaks="1 day", labels =label_date_short()) + 
  theme(axis.text.x=element_text(hjust=1)) +
    
  # Add label
  labs(title = "Locations by Car IDs - Hang out At Museum and Park", y = "Car ID", x = "Date", color = "Locations") 

# Make gantt chart interactive
ggplotly(events_timeline_park, tooltip = c("text"))
```

#### 5.3.10 Informal relationships between individuals through meals

```{r food events timeline}
# Change data into long form for line chart
events_matched_long_food <- events_matched %>% 
  filter(category == "Food") %>%
  mutate(rownum = row_number()) %>%
  dplyr::select(!event_stop) %>%
  pivot_longer(cols = starts_with("event_st"), names_to = "event_start_stop", values_to = "timestamp") 

# Plot timeline data as gantt chart
events_timeline_food <- ggplot(events_matched_long_food, 
                          aes(text = paste('id:', id,
                                           '<br>Name:', name,
                                           '<br>Department:', department,
                                           '<br>Tite:', title,
                                           '<br>Timestamp:', timestamp,
                                           '<br>Location: ', location,
                                           '<br>Category: ', category))) +
  # Add line range
  geom_line(mapping=aes(y=id,x=timestamp,color=location,group=rownum),size=1) +
  
  # Change theme
  theme_minimal() + 
  
  # Modify scale
  scale_y_discrete(limits = rev) +
  scale_x_datetime(date_breaks="1 day", labels =label_date_short()) + 
  theme(axis.text.x=element_text(hjust=1)) +
    
  # Add label
  labs(title = "Locations by Car IDs", y = "Car ID", x = "Date", color = "Locations") 

# Make gantt chart interactive
ggplotly(events_timeline_food, tooltip = c("text"))
```

```{r food events - brewed awakenings timeline}
# Change data into long form for line chart
events_matched_long_BA <- events_matched %>% 
  filter(location == "Brewed Awakenings") %>%
  mutate(rownum = row_number()) %>%
  dplyr::select(!event_stop) %>%
  pivot_longer(cols = starts_with("event_st"), names_to = "event_start_stop", values_to = "timestamp") 

# Plot timeline data as gantt chart
events_timeline_BA <- ggplot(events_matched_long_BA, 
                          aes(text = paste('id:', id,
                                           '<br>Name:', name,
                                           '<br>Department:', department,
                                           '<br>Tite:', title,
                                           '<br>Timestamp:', timestamp,
                                           '<br>Location: ', location,
                                           '<br>Category: ', category))) +
  # Add line range
  geom_line(mapping=aes(y=id,x=timestamp,color=location,group=rownum),size=1) +
  
  # Change theme
  theme_minimal() + 
  
  # Modify scale
  scale_y_discrete(limits = rev) +
  scale_x_datetime(date_breaks="1 day", labels =label_date_short()) + 
  theme(axis.text.x=element_text(hjust=1)) +
    
  # Add label
  labs(title = "Locations by Car IDs - Surveillance at Brewed Awakenings", y = "Car ID", x = "Date", color = "Locations") 

# Make gantt chart interactive
ggplotly(events_timeline_BA, tooltip = c("text"))
```







#### 5.3.11 Match Car IDs with credit card and loyalty card IDs

```{r match gps events with spend events}
# Fuzzy match all events with credit card and loyalty card transactions to find unique pairs of car ID and credit card & loyalty card IDs
spend_match <- trans_final %>%
  fuzzy_left_join(events_matched, 
                  by = c("timestamp" = "event_start_time", 
                        "timestamp" = "event_stop_time", "location" = "location"),
                  match_fun = list(`>`,`<`,`==`)) %>% 
  dplyr::select(name, department, title, id, last4ccnum, loyaltynum) %>%
  group_by(name, department, title, id, last4ccnum, loyaltynum) %>% 
  summarise(n = n()) %>%
  ungroup() %>%
  drop_na(n) %>%
  group_by(last4ccnum) %>%
  drop_na(id) %>%
  arrange(desc(n)) %>%
  slice_head() %>%
  ungroup()

# Each car driver assume to have 1 primary credit card
spend_match_2 <- spend_match %>% 
  filter(!is.na(name)) %>%
  group_by(id) %>%
  arrange(desc(n)) %>%
  slice_head() %>%
  ungroup()

# Each of 9 truck drivers assume to have 1 credit card based on the spend pattern we saw earlier 
car_spend_match <- spend_match %>% 
  filter(is.na(name)) %>%
  rbind(spend_match_2)

# Ownership of Credit Card and Car ID
DT::datatable(car_spend_match, filter = 'top')
```



#### 5.3.12 Plot timeline of events with spending transactions

From the interactive timeline charts of the GPS locations and spend locations, we observe that employees usually make a corresponding spend on their credit cards when visiting a location with some exceptions. 
```{r spend gps events fixed postings timeline}

trans_final_id <- car_spend_match %>%
  left_join(trans_final, by = c("last4ccnum")) %>%
  filter(!is.na(id))

# Plot timeline data as gantt chart
g1 <- ggplot() +
  # Add line 
  geom_line(events_matched_long, mapping=aes(y=id,x=timestamp,color=location,group=rownum, 
                                             text = paste('id:', id,
                                           '<br>Name:', name,
                                           '<br>Department:', department,
                                           '<br>Title:', title,
                                           '<br>Timestamp:', timestamp,
                                           '<br>Location: ', location,
                                           '<br>Category: ', category)),size=1) +
  
  # Change theme
  theme_minimal() + 
  
  # Modify scale
  scale_y_discrete(limits = rev) +
  scale_x_datetime(date_breaks="1 day", labels =label_date_short()) + 
  theme(axis.text.x=element_text(hjust=1)) 
    

# Plot timeline data as gantt chart
g2 <- ggplot() +
  # Add point 
  geom_point(trans_final_id,mapping=aes(y=id, x=timestamp, color=location,
                                        text = paste('id:', id,
                                           '<br>Name:', name,
                                           '<br>Department:', department,
                                           '<br>Title:', title,
                                           '<br>Last 4 Credit Card Num:', last4ccnum,
                                           '<br>Loyalty Num:', loyaltynum.y,
                                           '<br>Price: $', price,
                                           '<br>Timestamp:', timestamp,
                                           '<br>Location: ', location)),size=1) +
  
  # Change theme
  theme_minimal() + 
  
  # Modify scale
  scale_y_discrete(limits = rev) +
  scale_x_datetime(date_breaks="1 day", labels =label_date_short()) + 
  theme(axis.text.x=element_text(hjust=1)) +
    
  # Add label
  labs(title = "GPS and Spend Transactions by Locations and ID", y = "Car ID", x = "Date", color = "Locations") 

  
# Make gantt chart interactive
subplot(ggplotly(g1, tooltip = c("text")), ggplotly(g2, tooltip = c("text")))
```

Find transactions that do not match with GPS locations and GPS locations that did not register any spend
```{r spend gps events discrepancies}

# Events with timestamp
trans_final_id <- car_spend_match %>%
  left_join(trans_final, by = c("last4ccnum")) %>%
  filter(!is.na(id)) %>%
  dplyr::select(timestamp, datestamp, datestamp_loyalty, location, id, last4ccnum, loyaltynum.y, price)

# GPS events at POIs
POI_events_only <- events_matched %>% 
  filter(!category %in% c("Home","Unknown")) %>% 
  filter(!location %in% c("GASTech","Kronos Capitol"))

# Fuzzy match all events with credit card and loyalty card transactions to find unique pairs of car ID and credit card & loyalty card IDs
spend_match <- trans_final_id %>%
  fuzzy_full_join(POI_events_only, 
                  by = c("timestamp" = "event_start_time", 
                        "timestamp" = "event_stop_time", 
                        "location" = "location",
                        "id" = "id"),
                  match_fun = list(`>`,`<`,`==`,`==`))

spend_mismatch <- spend_match %>%
  filter(is.na(timestamp)|is.na(event_start_time))

# Change data into long form for line chart
gps_mismatched_long <- spend_mismatch %>% 
  filter(!is.na(event_start_time)) %>% 
  rename(id = id.y, location = location.y) %>% 
  dplyr::select(event_date, event_start_time, event_stop_time, diff, lat, long, geohash, id, name, department, title, location, category) %>%
  mutate(rownum = row_number()) %>%
  pivot_longer(cols = starts_with("event_st"), names_to = "event_start_stop", values_to = "timestamp")

spend_mismatch_prep <- spend_mismatch %>%
  filter(!is.na(timestamp)) %>%
  rename(id = id.x, location = location.x) %>%
  dplyr::select(timestamp, datestamp, datestamp_loyalty, location, id, last4ccnum, loyaltynum.y, price)
                  
# Plot timeline data as gantt chart
events_timeline_mismatch <- ggplot() +
  # Add line & point
  geom_line(gps_mismatched_long, mapping=aes(y=id,x=timestamp,color=location,group=rownum, 
                                             text = paste('id:', id,
                                           '<br>Name:', name,
                                           '<br>Department:', department,
                                           '<br>Title:', title,
                                           '<br>Timestamp:', timestamp,
                                           '<br>Location: ', location,
                                           '<br>Category: ', category)),size=1) +
  geom_point(spend_mismatch_prep, mapping=aes(y=id, x=timestamp, fill=location,
                                        text = paste('id:', id,
                                           '<br>Last 4 Credit Card Num:', last4ccnum,
                                           '<br>Loyalty Num:', loyaltynum.y,
                                           '<br>Price: $', price,
                                           '<br>Timestamp:', timestamp,
                                           '<br>Location: ', location)),size=4, alpha=0.3) +
  
  # Change theme
  theme_minimal() + 
  
  # Modify scale
  scale_y_discrete(limits = rev) +
  scale_x_datetime(date_breaks="1 day", labels =label_date_short()) + 
  theme(axis.text.x=element_text(hjust=1)) +
    
  # Add label
  labs(title = "Mismatch Locations", y = "Car ID", x = "Date", color = "Locations", fill = "Locations") 
  
# Make gantt chart interactive
ggplotly(events_timeline_mismatch, tooltip = c("text"))                  

```


Kronos Mart credit card postings were posted approximately 12 hours after the purchase was made.  
```{r spend + gps events  - Kronos Mart timeline}
# Change data into long form for line chart
trans_final_KM <- car_spend_match %>%
  left_join(trans_final, by = c("last4ccnum")) %>%
  filter(location == "Kronos Mart")

events_matched_long_KM <- events_matched_long %>%
  filter(location == "Kronos Mart")

# Plot timeline data as gantt chart
events_timeline_spend_KM <- ggplot() +
  # Add line & point
  geom_line(events_matched_long_KM, mapping=aes(y=id,x=timestamp,color=location,group=rownum, 
                                             text = paste('id:', id,
                                           '<br>Name:', name,
                                           '<br>Department:', department,
                                           '<br>Title:', title,
                                           '<br>Timestamp:', timestamp,
                                           '<br>Location: ', location,
                                           '<br>Category: ', category)),size=1) +
  geom_point(trans_final_KM,mapping=aes(y=id, x=timestamp, fill=location,
                                        text = paste('id:', id,
                                           '<br>Name:', name,
                                           '<br>Department:', department,
                                           '<br>Title:', title,
                                           '<br>Last 4 Credit Card Num:', last4ccnum,
                                           '<br>Loyalty Num:', loyaltynum.y,
                                           '<br>Price: $', price,
                                           '<br>Timestamp:', timestamp,
                                           '<br>Location: ', location)),size=4, alpha=0.3) +
  
  # Change theme
  theme_minimal() + 
  
  # Modify scale
  scale_y_discrete(limits = rev) +
  scale_x_datetime(date_breaks="1 day", labels =label_date_short()) + 
  theme(axis.text.x=element_text(hjust=1)) +
    
  # Add label
  labs(title = "Kronos Mart - Visits and Spend Transactions Mismatched", y = "Car ID", x = "Date", color = "GPS Locations - Line", fill = "Spend Locations - Circle") 

# Make gantt chart interactive
ggplotly(events_timeline_spend_KM, tooltip = c("text"))
```

Nils' (Car ID 1) credit card (9551) was used at Frydos Autosupply n More on 13 Jan at but he was not present at the location based on his GPS. Nils was actually at Ouzeri Elian. Instead, Minke (Car ID 24) was present at the shop. We can infer that Minke had stolen Nils' credit card for fraudulent use. 
```{r spend gps events Frydos Autosupply n More timeline}

# Change data into long form for line chart
trans_final_FA <- car_spend_match %>%
  left_join(trans_final, by = c("last4ccnum")) %>%
  filter(location == "Frydos Autosupply n' More")

events_matched_long_FA <- events_matched_long %>%
  filter(location == "Frydos Autosupply n' More")

# Plot timeline data as gantt chart
events_timeline_spend_FA <- ggplot() +
  # Add line & point
  geom_line(events_matched_long_FA, mapping=aes(y=id,x=timestamp,color=location,group=rownum, 
                                             text = paste('id:', id,
                                           '<br>Name:', name,
                                           '<br>Department:', department,
                                           '<br>Title:', title,
                                           '<br>Timestamp:', timestamp,
                                           '<br>Location: ', location,
                                           '<br>Category: ', category)),size=1) +
  geom_point(trans_final_FA,mapping=aes(y=id, x=timestamp, fill=location,
                                        text = paste('id:', id,
                                           '<br>Name:', name,
                                           '<br>Department:', department,
                                           '<br>Title:', title,
                                           '<br>Last 4 Credit Card Num:', last4ccnum,
                                           '<br>Loyalty Num:', loyaltynum.y,
                                           '<br>Price: $', price,
                                           '<br>Timestamp:', timestamp,
                                           '<br>Location: ', location)),size=4, alpha=0.3) +
  
  # Change theme
  theme_minimal() + 
  
  # Modify scale
  scale_y_discrete(limits = rev) +
  scale_x_datetime(date_breaks="1 day", labels =label_date_short()) + 
  theme(axis.text.x=element_text(hjust=1)) +
    
  # Add label
  labs(title = "Frydos Autosupply n' More - Visits and Spend Transactions by Car IDs", y = "Car ID", x = "Date", color = "GPS Locations - Line", fill = "Spend Locations - Circle") 
  
# Make gantt chart interactive
ggplotly(events_timeline_spend_FA, tooltip = c("text"))
```


Matching the GPS and spend transactions for the Food establishments with fixed credit card postings shows that the credit card postings are usually done on the same day of the purchase. 
```{r spend gps events food timeline}

# Change data into long form for line chart
trans_final_cafe <- car_spend_match %>%
  left_join(trans_final, by = c("last4ccnum")) %>%
  filter(location %in% c("Bean There Done That", "Brewed Awakenings","Jack's Magical Beans"))

events_matched_long_cafe <- events_matched_long %>%
  filter(location %in% c("Bean There Done That", "Brewed Awakenings","Jack's Magical Beans"))

# Plot timeline data as gantt chart
events_timeline_spend_cafe <- ggplot() +
  # Add line & point
  geom_line(events_matched_long_cafe, mapping=aes(y=id,x=timestamp,color=location,group=rownum, 
                                             text = paste('id:', id,
                                           '<br>Name:', name,
                                           '<br>Department:', department,
                                           '<br>Title:', title,
                                           '<br>Timestamp:', timestamp,
                                           '<br>Location: ', location,
                                           '<br>Category: ', category)),size=1) +
  geom_point(trans_final_cafe,mapping=aes(y=id, x=timestamp, fill=location,
                                        text = paste('id:', id,
                                           '<br>Name:', name,
                                           '<br>Department:', department,
                                           '<br>Title:', title,
                                           '<br>Last 4 Credit Card Num:', last4ccnum,
                                           '<br>Loyalty Num:', loyaltynum.y,
                                           '<br>Price: $', price,
                                           '<br>Timestamp:', timestamp,
                                           '<br>Location: ', location)),size=4, alpha=0.3) +
  
  # Change theme
  theme_minimal() + 
  
  # Modify scale
  scale_y_discrete(limits = rev) +
  scale_x_datetime(date_breaks="1 day", labels =label_date_short()) + 
  theme(axis.text.x=element_text(hjust=1)) +
    
  # Add label
  labs(title = "Cafes with Fixed Posting - Visits and Spend Transactions Mismatched", y = "Car ID", x = "Date", color = "GPS Locations - Line", fill = "Spend Locations - Circle") 
  
# Make gantt chart interactive
ggplotly(events_timeline_spend_cafe, tooltip = c("text"))
```

#### 5.3.13 Plot timeline of events of individuals

```{r individual event - 22 & 28 timeline}
# Change data into long form for line chart
events_matched_long_2 <- events_matched %>% 
  mutate(event_date_end = as_date(event_stop_time)) %>%
  mutate(flag = ifelse(event_date_end > event_date,1,0)) %>%
  filter(flag == 0) %>%
  filter(id %in% c("22","28")) %>%
  mutate(event_start_time = format(event_start_time, format = "%H:%M:%S")) %>%
  mutate(event_start_time = as.POSIXct(event_start_time, format = "%H:%M:%S")) %>%
  mutate(event_stop_time = format(event_stop_time, format = "%H:%M:%S")) %>%
  mutate(event_stop_time = as.POSIXct(event_stop_time, format = "%H:%M:%S")) %>%
  mutate(rownum = row_number()) %>%
  dplyr::select(!event_stop) %>%
  pivot_longer(cols = starts_with("event_st"), names_to = "event_start_stop", values_to = "timestamp")

trans_final_2 <- car_spend_match %>%
  left_join(trans_final, by = c("last4ccnum")) %>%
  filter(id %in% c("22","28")) %>%
  mutate(timestamp = format(timestamp, format = "%H:%M:%S")) %>%
  mutate(timestamp = as.POSIXct(timestamp, format = "%H:%M:%S"))

# Plot timeline data as gantt chart
events_timeline_2 <- ggplot() +
  # Add line range
  geom_line(events_matched_long_2, mapping=aes(y=event_date,x=timestamp,color=location,group=rownum, text = paste('id:', id,
                                           '<br>Name:', name,
                                           '<br>Department:', department,
                                           '<br>Tite:', title,
                                           '<br>Timestamp:', format(timestamp, format = "%H:%M"),
                                           '<br>Location: ', location,
                                           '<br>Category: ', category)),size=1) +
  geom_point(trans_final_2,mapping=aes(y=datestamp, x=timestamp, fill=location,
                                        text = paste('id:', id,
                                           '<br>Name:', name,
                                           '<br>Department:', department,
                                           '<br>Title:', title,
                                           '<br>Last 4 Credit Card Num:', last4ccnum,
                                           '<br>Loyalty Num:', loyaltynum.y,
                                           '<br>Price: $', price,
                                           '<br>Timestamp:', timestamp,
                                           '<br>Location: ', location)),size=4, alpha=0.3) +
  
  # Change theme
  theme_minimal() +

  # Modify scale
  scale_y_date(breaks="1 day", labels=label_date_short()) +
  scale_x_datetime(breaks="1 hour", labels=time_format(format = "%H:%M", tz = Sys.timezone(location=TRUE))) + 
  theme(axis.text.x=element_text(angle=45, hjust=1)) +

  facet_wrap(~id, ncol = 2) +

  # Add label
  labs(title = "Locations Visted By Kanon(22) and Borrasca(28)", y = "Event Start Date", x = "Time", color = "Locations") 

# Make gantt chart interactive
ggplotly(events_timeline_2, tooltip = c("text"))

```

Bertrand changed credit cards. 
```{r individual event - 29 timeline}
# Change data into long form for line chart
events_matched_long_3 <- events_matched %>% 
  mutate(event_date_end = as_date(event_stop_time)) %>%
  mutate(flag = ifelse(event_date_end > event_date,1,0)) %>%
  filter(flag == 0) %>%
  filter(id %in% c("29")) %>%
  mutate(event_start_time = format(event_start_time, format = "%H:%M:%S")) %>%
  mutate(event_start_time = as.POSIXct(event_start_time, format = "%H:%M:%S")) %>%
  mutate(event_stop_time = format(event_stop_time, format = "%H:%M:%S")) %>%
  mutate(event_stop_time = as.POSIXct(event_stop_time, format = "%H:%M:%S")) %>%
  mutate(rownum = row_number()) %>%
  dplyr::select(!event_stop) %>%
  pivot_longer(cols = starts_with("event_st"), names_to = "event_start_stop", values_to = "timestamp")

trans_final_3 <- car_spend_match %>%
  left_join(trans_final, by = c("last4ccnum")) %>%
  filter(id %in% c("29")) %>%
  mutate(timestamp = format(timestamp, format = "%H:%M:%S")) %>%
  mutate(timestamp = as.POSIXct(timestamp, format = "%H:%M:%S"))

# Plot timeline data as gantt chart
events_timeline_3 <- ggplot() +
  # Add line range
  geom_line(events_matched_long_3, mapping=aes(y=event_date,x=timestamp,color=location,group=rownum, text = paste('id:', id,
                                           '<br>Name:', name,
                                           '<br>Department:', department,
                                           '<br>Tite:', title,
                                           '<br>Timestamp:', format(timestamp, format = "%H:%M"),
                                           '<br>Location: ', location,
                                           '<br>Category: ', category)),size=1) +
  geom_point(trans_final_3,mapping=aes(y=datestamp, x=timestamp, fill=location,
                                        text = paste('id:', id,
                                           '<br>Name:', name,
                                           '<br>Department:', department,
                                           '<br>Title:', title,
                                           '<br>Last 4 Credit Card Num:', last4ccnum,
                                           '<br>Loyalty Num:', loyaltynum.y,
                                           '<br>Price: $', price,
                                           '<br>Timestamp:', timestamp,
                                           '<br>Location: ', location)),size=4, alpha=0.3) +
  
  # Change theme
  theme_minimal() +

  # Modify scale
  scale_y_date(breaks="1 day", labels=label_date_short()) +
  scale_x_datetime(breaks="1 hour", labels=time_format(format = "%H:%M", tz = Sys.timezone(location=TRUE))) + 
  theme(axis.text.x=element_text(angle=45, hjust=1)) 

  #facet_wrap(~id, ncol = 2) +

  # Add label
  labs(title = "Locations Visted By Bertrand (29)", y = "Event Start Date", x = "Time", color = "Locations") 

# Make gantt chart interactive
ggplotly(events_timeline_3, tooltip = c("text"))

```

Find transactions that do not match with GPS locations and GPS locations that did not register any spend
```{r spend gps events remaining discrepancies}

# GPS events at POIs
POI_events_only_2 <- events_matched %>% 
  filter(!category %in% c("Home","Unknown")) %>% 
  filter(!location %in% c("GASTech","Kronos Capitol","Kronos Mart","Bean There Done That","Brewed Awakenings","Jack's Magical Beans")) %>%
  filter(!id %in% c("22","28","29"))

# Fuzzy match all events with credit card and loyalty card transactions to find unique pairs of car ID and credit card & loyalty card IDs
spend_match_2 <- trans_final_id %>%
  fuzzy_full_join(POI_events_only, 
                  by = c("timestamp" = "event_start_time", 
                        "timestamp" = "event_stop_time", 
                        "location" = "location",
                        "id" = "id"),
                  match_fun = list(`>`,`<`,`==`,`==`))

spend_mismatch <- spend_match %>%
  filter(is.na(timestamp)|is.na(event_start_time))

# Change data into long form for line chart
gps_mismatched_long_2 <- spend_mismatch_2 %>% 
  filter(!is.na(event_start_time)) %>% 
  rename(id = id.y, location = location.y) %>% 
  dplyr::select(event_date, event_start_time, event_stop_time, diff, lat, long, geohash, id, name, department, title, location, category) %>%
  mutate(rownum = row_number()) %>%
  pivot_longer(cols = starts_with("event_st"), names_to = "event_start_stop", values_to = "timestamp")

spend_mismatch_prep_2 <- spend_mismatch_2 %>%
  filter(!is.na(timestamp)) %>%
  rename(id = id.x, location = location.x) %>%
  dplyr::select(timestamp, datestamp, datestamp_loyalty, location, id, last4ccnum, loyaltynum.y, price)
                  
# Plot timeline data as gantt chart
events_timeline_mismatch_2 <- ggplot() +
  # Add line & point
  geom_line(gps_mismatched_long_2, mapping=aes(y=id,x=timestamp,color=location,group=rownum, 
                                             text = paste('id:', id,
                                           '<br>Name:', name,
                                           '<br>Department:', department,
                                           '<br>Title:', title,
                                           '<br>Timestamp:', timestamp,
                                           '<br>Location: ', location,
                                           '<br>Category: ', category)),size=1) +
  geom_point(spend_mismatch_prep_2, mapping=aes(y=id, x=timestamp, fill=location,
                                        text = paste('id:', id,
                                           '<br>Last 4 Credit Card Num:', last4ccnum,
                                           '<br>Loyalty Num:', loyaltynum.y,
                                           '<br>Price: $', price,
                                           '<br>Timestamp:', timestamp,
                                           '<br>Location: ', location)),size=4, alpha=0.3) +
  
  # Change theme
  theme_minimal() + 
  
  # Modify scale
  scale_y_discrete(limits = rev) +
  scale_x_datetime(date_breaks="1 day", labels =label_date_short()) + 
  theme(axis.text.x=element_text(hjust=1)) +
    
  # Add label
  labs(title = "Mismatch Locations", y = "Car ID", x = "Date", color = "Locations", fill = "Locations") 
  
# Make gantt chart interactive
ggplotly(events_timeline_mismatch_2, tooltip = c("text"))                  

```



#### 5.4 Suspicious Individuals

Isolating individuals who visited unknown locations and was carrying out surveillance on Executives houses.
```{r individual event timeline}
# Change data into long form for line chart
events_matched_long_1 <- events_matched %>% 
  mutate(event_date_end = as_date(event_stop_time)) %>%
  mutate(flag = ifelse(event_date_end > event_date,1,0)) %>%
  filter(flag == 0) %>%
  filter(id %in% c("13","15","16","21","24")) %>%
  mutate(event_start_time = format(event_start_time, format = "%H:%M:%S")) %>%
  mutate(event_start_time = as.POSIXct(event_start_time, format = "%H:%M:%S")) %>%
  mutate(event_stop_time = format(event_stop_time, format = "%H:%M:%S")) %>%
  mutate(event_stop_time = as.POSIXct(event_stop_time, format = "%H:%M:%S")) %>%
  mutate(rownum = row_number()) %>%
  dplyr::select(!event_stop) %>%
  pivot_longer(cols = starts_with("event_st"), names_to = "event_start_stop", values_to = "timestamp")


trans_final_1 <- car_spend_match %>%
  left_join(trans_final, by = c("last4ccnum")) %>%
  filter(id %in% c("13","15","16","21","24")) %>%
  mutate(timestamp = format(timestamp, format = "%H:%M:%S")) %>%
  mutate(timestamp = as.POSIXct(timestamp, format = "%H:%M:%S"))

# Plot timeline data as gantt chart
events_timeline_1 <- ggplot(events_matched_long_1, 
                          aes(text = paste('id:', id,
                                           '<br>Name:', name,
                                           '<br>Department:', department,
                                           '<br>Tite:', title,
                                           '<br>Timestamp:', format(timestamp, format = "%H:%M"),
                                           '<br>Location: ', location,
                                           '<br>Category: ', category))) +
  # Add line & point
  geom_line(mapping=aes(y=event_date,x=timestamp,color=location,group=rownum),size=1) +
  
  geom_point(trans_final_1,mapping=aes(y=datestamp, x=timestamp, fill=location,
                                        text = paste('id:', id,
                                           '<br>Name:', name,
                                           '<br>Department:', department,
                                           '<br>Title:', title,
                                           '<br>Last 4 Credit Card Num:', last4ccnum,
                                           '<br>Loyalty Num:', loyaltynum.y,
                                           '<br>Price: $', price,
                                           '<br>Timestamp:', timestamp,
                                           '<br>Location: ', location)),size=4, alpha=0.3) +
  
  # Change theme
  theme_minimal() +

  # Modify scale
  scale_y_date(breaks="1 day", labels=label_date_short()) +
  scale_x_datetime(breaks="1 hour", labels=time_format(format = "%H:%M", tz = Sys.timezone(location=TRUE))) + 
  theme(axis.text.x=element_text(angle=45, hjust=1)) +

  facet_wrap(~id, ncol = 2) +

  # Add label
  labs(title = "Locations Visted By Car IDs of Suspects", y = "Event Start Date", x = "Time", color = "Locations") 

# Make gantt chart interactive
ggplotly(events_timeline_1, tooltip = c("text"))

```

#### 5.3.14 Link timeline to map to find suspicious locations

#### Prep data for timeline linking to map
```{r}
map <- ggplot(location_labels, aes(long, lat, text = paste("geohash:", geohash))) + 
      
    theme(panel.border = element_blank(),
    panel.grid.major = element_blank(), 
    panel.grid.minor = element_blank(), 
    panel.background = element_blank(),
    axis.line = element_line(colour = "grey")) +
  
  # Add paths
    geom_line(gps_path, aes())

  # Add points
    geom_point(aes(color=category, alpha = 0.3)) +
    geom_text(aes(label =location), size=2)

# Use ggplotly to make it interactive so we can pick the right geohash visually
pp <- ggplotly(map)
layout(pp, images = list(
  list(
    source = "https://raw.githubusercontent.com/kpokp/ISSS608_blog/main/_posts/2021-06-10-individual-assignment/images/MC2-tourist.jpg",
    opacity = 0.3,
    layer = "below",
    xref = "x",
    yref = "y",
    sizing="stretch",
    x= 24.8244,
    y= 36.0952,
    sizex= 0.0852,
    sizey= 0.05
  )
))
```

#### Prep data for timeline linking to map
```{r}

location_labels_sf <- st_as_sf(location_labels, coords = c("long", "lat"), crs = 4326) 
gps_path_selected <- gps_path %>% filter(id == c("1"))

tmap_mode("view")

tm_shape(ap) +
  tm_rgb(ap, r=1,g=2,b=3,
         alpha = NA,
         saturation = 1,
         interpolate = TRUE,
         max.value = 255) +
tm_shape(gps_path_selected) +
  tm_lines() +
tm_shape(location_labels_sf) +
  tm_squares(col = "category",
          id = "location", 
          popup.vars = "Location Type:" =="category",
          size = 0.1)
```








## Insights and Conlusions

### Question 1 
Using just the credit and loyalty card data, identify the most popular locations, and when they are popular. What anomalies do you see? What corrections would you recommend to correct these anomalies? Please limit your answer to 8 images and 300 words. 

#### Q1 Most popular locations
The top 3 popular locations belong to the Food categories, such as Katerina's Cafe, Brew've Been Served and Hippokampos. Katerina's Cafe peaks on Saturday, Hippokampus is busier on weekdays than weekends and Brew've been Served is closed on weekends.
```{r Q1 popular locations, echo=FALSE}
heatmap_spend_gg
```

#### Q1 Anomaly - Credit Card Posting at Fixed Hour
Some locations only have credit card transactions posted in 1 hour within the day e.g. Bean There Done That and Brewed Awakenings, Jack's Magical Beans, Coffee Shack at 12PM.
Corrections: Correct posting timing if GPS record shows otherwise.
```{r Q1 credit card postings, echo=FALSE}
heatmap_cc_gg
```

#### Q1 Anomaly - Kronos Mart Post Credit Card Record 1 Day Late
Kronos Mart's credit card postings usually come 1 day late, which result in strange timing of postings at 3-9AM. 
Corrections: Correct posting timing based on available GPS record or take the date of loyalty card record.
```{r Q1 late postings, echo=FALSE}
kronos_delay <- transact_match_3 %>%
  mutate(CreditCard = as_date(timestamp.x)) %>%
  dplyr::select(price, location, CreditCard, timestamp.y) %>% 
  rename(LoyaltyCard = timestamp.y) %>%
  pivot_longer(!c(location,price), names_to = "type", values_to = "date")
        
# Plot scatterplot of Kronos Mart delayed postings
p <- ggplot(kronos_delay, aes(x=date, y = price, color = type, label = price)) +
  geom_point() +
  geom_text(check_overlap = TRUE, size = 2, nudge_x = 0.5) +
  theme_minimal() +
  labs(title = "Daily Transactions of Kronos Mart by Card Type",
                   x = "Date", y = "Price ($)", color = "Card Type")

ggplotly(p)
```

#### Q1 Anomaly - Cashbacks in Multiples of $20
Some loyalty card transactions have prices lower than the corresponding credit card transactions. The denominations are usually in multiples of $20, ranging from $20 to $80. Locations vary widely, suggesting a loyalty cashback scheme. 
Corrections: Use credit card amount as loyalty price is non-consequential to investigation.
```{r Q1 cashbacks, echo=FALSE}
# 219 transactions match with price range with price difference in increments of $20.
p <- ggplot(transact_match_2, 
            aes(x = price_diff, fill = location.x, 
                text = paste('Price Difference: $', price_diff,
                         '<br>Location: ', location.x,
                         '<br>Count: ', n))) + 
  geom_histogram(boundary = 1, position = "stack") +
  labs(title = "Distribution of Cashback by Locations",
                   x = "Cashback Amount ($)", y = "Count", fill = "Locations") +
  theme_minimal()
  
ggplotly(p, tooltip = c("text"))  
```

#### Q1 Anomaly - Multiple Credit Cards to Loyalty Cards
Unexpected relationships between credit card 1286 and loyalty card L3288, and loyalty card L6267 is associated with credit cards 6691 & 6899.
Corrections: Care was taken to join credit card and loyalty card transactions to prevent duplication.
```{r multiple match, echo=FALSE}
netwkgraph
```

#### Q1 Anomaly - High Spend Due to Possible Fraud
The highest spend on the credit card is $10k on Frydos Autospply n' More that does not have a loyalty card transaction, belonging to credit card 9551. 9 credit cards register many high spends that are at industrial areas tying to the 9 truck drivers. The other highest spend that is not in industrial location is on Albert's Fine Clothing by credit card 1321. 

Using the data table, we noticed that there is a gap between the the $10k spend on 13 Jan and the next purchase by credit card 9551 on 16 Jan. This suggests that the high spend might be due to fraud. 
```{r high spends, echo=FALSE}
crosstalk::bscols(gg_cc, DT::datatable(d), widths = 10)
```


#### Q1 Anomaly - Sparse Spends
```{r sparse spends}
ly_count <- trans_all %>%
  group_by(datestamp, last4ccnum) %>%
  summarise(n = n()) %>% 
  drop_na() %>% ungroup()

ly_count_p <- ggplot(ly_count, aes(x = datestamp, y = n)) + 
  geom_bar(stat='identity') +
  facet_wrap(~last4ccnum) 

ggplotly(ly_count_p)
```



### Question 2
Add the vehicle data to your analysis of the credit and loyalty card data. How does your assessment of the anomalies in question 1 change based on this new data? What discrepancies between vehicle, credit, and loyalty card data do you find? Please limit your answer to 8 images and 500 words.


#### Q2 Anomaly - Confirm Credit Card Posting at Fixed Hour was Wrong
Bean There Done That, Brewed Awakenings and Jack's Magic Beans only operate in weekday mornings and posted credit card transactions at 12 noon.
```{r credit card posting, echo=FALSE}
```

#### Q2 Anomaly - Confirm Kronos Mart Posting was Wrong. 
Kronos Mart's credit card postings were posted around 12 hours after purchase. 
```{r Kronos Mart posting, echo=FALSE}
```

#### Q2 Anomaly - Confirm High Spend due to Fraud
Minke(24) likely stole Lars(1) credit card to for high spend on Frydos Autosupply n 'More. 
```{r high spends, echo=FALSE}
```




### Question 3
Can you infer the owners of each credit card and loyalty card? What is your evidence? Where are there uncertainties in your method? Where are there uncertainties in the data? Please limit your answer to 8 images and 500 words.

#### Q3 Owners of Credit Cards and Loyalty Card
```{r gps & credit card match, echo=FALSE}
```

#### Q3 Evidence - Credit Card Spends without Location Corroborated
```{r gps & credit card mismatch, echo=FALSE}
```

#### Q3 Uncertainty - Locations Close to Each Other
```{r uncertainty of match, echo=FALSE}
```



### Question 4
Given the data sources provided, identify potential informal or unofficial relationships among GASTech personnel. Provide evidence for these relationships. Please limit your response to 8 images and 500 words.

#### Q4 Relationships - Inga(13), Isia(16), Loreto(15), Hennie(21), Minke(24) conducting suspicious activities together
```{r relationship 1, echo=FALSE}
```

#### Q4 Relationships - Elsa(7) and Brand(33) have rendezvous middle of workdays
```{r relationship 2, echo=FALSE}
```

#### Q4 Relationships - Kanon(22) and Borrasca(28) are seeing each other
```{r relationship 3, echo=FALSE}
```

#### Q4 Relationships - Dinner party among employees from Engineers and IT departments
```{r relationship 4, echo=FALSE}
```

#### Q4 Relationships - Shared homes of employees
```{r relationship 5, echo=FALSE}
```

#### Q4 Relationships - Hennie's (21) second home
```{r relationship 6, echo=FALSE}
```



### Question 5
Do you see evidence of suspicious activity? Identify 1- 10 locations where you believe the suspicious activity is occurring, and why. Please limit your response to 10 images and 500 words.


#### Q5 Location - Drives to unknown houses
```{r location 1, echo=FALSE}
```

#### Q5 Location - Surveillance on Executive houses
```{r location 2, echo=FALSE}
```

#### Q5 Location - Surveillance at Executive lunch spots
```{r location 3, echo=FALSE}
```

#### Q5 Location - Hennie's unusual home stays
```{r location 4, echo=FALSE}
```

#### Q5 Location - Bertrand's late night trip
```{r location 5, echo=FALSE}
```

