---
title: "TIL: Week 8 In-Class Exercise"
description: |
   Today I learnt in class: We start learning about interactive visualization using R. Tidyverse will be used to create the data science workflow. 
author:
  - name: Kelly Koh
    url: https://www.linkedin.com/in/kellykkw/
    affiliation: School of Computing and Information Systems, Singapore Management University
    affiliation_url: https://scis.smu.edu.sg/
preview: images/__.png
date: 06-26-2021
output:
  distill::distill_article:
    toc: true
    toc_depth: 3
    toc_float: true
    self_contained: false
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Install the necessary packages if they are not in library
```{r echo=TRUE, eval=TRUE, message=FALSE, warning=FALSE}
packages = c('DT','ggiraph','plotly','tidyverse')

for (p in packages) {
  if(!require(p, character.only = T)){
    install.packages(p)
  }
  library(p, character.only = T)
}
```
### Import data from csv and preview data
```{r echo=TRUE, eval=TRUE, message=FALSE, warning=FALSE}
exam_data <- read_csv("data/Exam_data.csv")

head(exam_data,10)
summary(exam_data)
```

The dataset is made up of 7 variables with 322 observations:

- `ID` is the unique key for each student
- `CLASS`, `GENDER`, `RACE` are categorical variables expressed as character
- `ENGLISH`, `MATHS`, `SCIENCE` are continuous variables defined as numeric values with decimal points 

### Plot histogram of the Math scores by Gender
```{r echo=TRUE, eval=TRUE, message=FALSE, warning=FALSE}
math_hist <- ggplot(data = exam_data, aes(x = MATHS, fill = GENDER)) +
  geom_histogram(bins = 10,
                 boundary = 100,
                 color = "black",
                 position = "stack"
                 ) +
  scale_x_continuous("MATHS", limits = c(0,100), breaks = seq(0,100,10)) +
  ggtitle("Distribution of Math Scores")

ggplotly(math_hist)

```

### Plot bar chart of race
```{r echo=TRUE, eval=TRUE, message=FALSE, warning=FALSE}
race_bar <- ggplot(data = exam_data, aes(x = RACE)) +
  geom_bar() +
  ggtitle("Distribution of Race")

ggplotly(race_bar)

```

### Plot dotplot of Math Scores by Race
```{r echo=TRUE, eval=TRUE, message=FALSE, warning=FALSE}
math_dot <- ggplot(data = exam_data, 
                   aes(x = MATHS, fill = RACE)) +
  geom_dotplot(binwidth = 2.5,
               dotsize = 0.5,
               stackgroups = TRUE) +
  scale_y_continuous(NULL, 
                     breaks = NULL) +
  ggtitle("Distribution of Math Scores")

math_dot

```